---
title: "StanModel"
output: html_document
---

# Setup
```{r}
pacman::p_load(
  tidyverse,    # Data manipulation and plotting
  here,         # Robust file paths
  cmdstanr,     # Stan interface
  rstan,
  posterior,    # Posterior analysis
  bayesplot,    # Plotting MCMC
  furrr,        # Parallel processing
  future,        # Future backend
  patchwork
)
```

```{r}
# Parallel Setup: Leave 2 cores free for system stability
n_cores <- parallel::detectCores() - 2
plan(multisession, workers = n_cores)
```

```{r}
# Optional: Set a default ggplot theme
theme_set(theme_classic())
```

# Simulating Data
```{r}
# Flag to control whether to regenerate simulation/fitting results for the knitted files
# Set to TRUE to rerun everything (takes time!), FALSE to load saved results.
regenerate_simulations <- TRUE # Or TRUE

```

```{r}
# --- Simulate Data for Fitting ---
# Goal: Generate data from known models/parameters to test our fitting procedures.
trials <- 120 # Number of trials per simulated agent
```


```{r}
RandomAgentNoise_f <- function(rate_prob, noise) {
  # Ensure rate is probability (in case log-odds are passed, though not here)
  choice <- rbinom(1, 1, rate_prob) # Base choice
  if (noise > 0 & rbinom(1, 1, noise) == 1) { # Check if noise occurs
    choice <- rbinom(1, 1, 0.5) # Override with random 50/50 choice
  }
  return(choice)
}
```

```{r}
# Define Agent - what kind of player is this simulated data from?
rl_agent_f <- function(opp_choice, estimated_value, alpha =0.4, tau=1) { 
  
  PE = opp_choice - estimated_value;
  estimated_value = estimated_value + (alpha * PE);
    
  # softmax
  p = exp(tau * estimated_value) / (exp(tau * estimated_value) + exp(tau * (1-estimated_value)));

  # choice based on softmax probabilities - prob needs to be a vector as it is for both hands
  choice <- sample(c(0, 1), 1, prob = c(1 - p, p))
  
   return(list(
    choice = choice,
    estimated_value = estimated_value
  ))
}
```


```{r} 
full_simulation <- function(alpha, tau) {

# Create empty lists for choices and feedback
RA_choices <- rep(NA, trials)
RL_choices <- rep(NA, trials)
feedback_RA <- rep(NA, trials)
feedback_RL <- rep(NA, trials)
estimated_value_tracked <- rep(NA, trials) 
estimated_value_tracked[1] <- 0.5 #tracking "left" hand - estiamted_value[1]
 

RA_choices[1] <- sample(c(0, 1), 1) #choice on trial 1 for agent 1 
RL_choices[1] <- sample(c(0, 1), 1) #choice on trial 1 for agent 2 

for (t in 2:trials) {
  # Feedback for agent 1 (RL) from previous trial (wins if match)
  prev_feedback1 <- ifelse(RL_choices[t - 1] == RA_choices[t - 1], 1, 0)
  feedback_RL[t - 1] <- prev_feedback1

  # Feedback for agent 2 (RA - Opponent) is the opposite (wins if mismatch)
  prev_feedback2 <- 1 - prev_feedback1
  feedback_RA[t - 1] <- prev_feedback2

  # Both agents choose based on their own previous feedback
  RA_choices[t] <- RandomAgentNoise_f(rate_prob=0.8, noise=0.2) #RA agent choice
  
  # RL agent choice and update estimated value based on previous feedback
  result <- rl_agent_f(
    opp_choice = RA_choices[t - 1], # opponents choice to drive the PE in the rl_agent_f
    estimated_value = estimated_value_tracked[t - 1],
    alpha=alpha,
    tau = tau
  )
  
  # RL agent choice and update estimated value defined for current trial
  RL_choices[t] <- result$choice
  estimated_value_tracked[t] <- result$estimated_value
}

# Defining the feedback across trials
feedback_RL[trials] <- ifelse(RL_choices[trials] == RA_choices[trials], 1, 0)
feedback_RA[trials] <- ifelse(RL_choices[trials] == RA_choices[trials], 0, 1)

# Create dataframe
df_RA_RL <- tibble(
  trial = 1:trials,
  Choice_RL = RL_choices,
  Opponent_RA = RA_choices,
  Feedback_RL = feedback_RL, # 1 if agent_RL won
  Feedback_RA = feedback_RA, # 1 if opponent_RA won
  RL_estimated_value_left = estimated_value_tracked,
  RL_estimated_value_right = 1-estimated_value_tracked,
  true_alpha=alpha,
  true_tau= tau
) %>% 
  group_by(true_alpha, true_tau) %>%
  mutate(cumulative_rate = cumsum(Choice_RL) / row_number()) %>%
  ungroup()

return(df_RA_RL)
}
```


```{r}
# --- Generate Data Across Different Parameters ---
# Define relative paths for saving/loading simulation data
sim_data_dir <- "Assignment02/data" # saves to data 'data' subdirectory
sim_data_file <- here(sim_data_dir, "Simulation_01.csv") 

# Create directory if it doesn't exist
if (!dir.exists(sim_data_dir)) dir.create(sim_data_dir, recursive = TRUE)

if (regenerate_simulations || !file.exists(sim_data_file)) {
  cat("Generating new simulation data...\n")
  # Use expand_grid for cleaner parameter combinations
  param_grid <- expand_grid(
    alpha = seq(0, 1, 0.1), # alpha levels
    tau = c(0.5, 1, 5, 10, 15, 20)     # Tau temp levels
  )

  
  # Use map_dfr for efficient simulation across parameters
  d <- pmap_dfr(param_grid, full_simulation) 

  # Save the generated data
  write_csv(d, sim_data_file)
  cat("Generated new simulation data and saved to", sim_data_file, "\n")
} else {
  # Load existing simulation data
  d <- read_csv(sim_data_file)
  cat("Loaded existing simulation data from", sim_data_file, "\n")
}
```

```{r}
# --- Visualize Simulated Data ---
# Plot cumulative alpha, faceted by tau level
p1 <- ggplot(d, aes(x = trial, y = cumulative_rate, group = true_tau, color = true_tau)) +
  geom_line(alpha = 0.8) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  scale_color_viridis_c() + # Use perceptually uniform color scale
  facet_wrap(~paste("Alpha =", true_alpha)) + # Facet by noise level
  labs(
    title = "Simulated Choice Behavior (Cumulative Rate)",
    subtitle = "Agents with different alpha and tau levels",
    x = "Trial Number",
    y = "Cumulative Proportion Choosing 'Right'",
    color = "True Tau Rate"
  ) +
  theme_minimal() +
  ylim(0, 1)

print(p1)

```
```{r}
# isolate for one alpha tau combination parameters then fit to the stan model

# Subset data for one specific agent simulation (alpha=0.8, tau=0.5)
d1 <- d %>% filter(true_alpha == 0.8, true_tau == 1) 

# Stan requires data in a list format
data_for_stan <- list(
  players = 1,
  n = trials,     # Number of trials (integer)
  choice = d1$Choice_RL,   # Vector of choices (0s and 1s) for the RL agent
  opp_choice = d1$Opponent_RA # Vector of opponent choices (0s and 1s) for the RA agent
)

str(data_for_stan) # Show the structure of the list

```


# Then need R framework to create the data, fit it to Stan, and then extract the posterior draws and visualise them.

#RUN STAN MODEL feeding it simulated data from above 
```{r}
## --- Compile and Fit ---
# Specify model file path
stan_file_bernoulli <- here("Assignment02/src/Stan_Model.stan")
# Define path for saving fitted model object
stan_results_dir <- "Assignment02/src/Stan_results"
model_file_bernoulli <- here(stan_results_dir, "04_SimpleBernoulli.rds")

if (!dir.exists(stan_results_dir)) dir.create(stan_results_dir)

if (regenerate_simulations || !file.exists(model_file_bernoulli)) {
  cat("Compiling and fitting Stan model (Bernoulli)...\n")
  # Compile the Stan model (only needs to be done once unless code changes)
  mod_bernoulli <- cmdstan_model(stan_file_bernoulli,
                                 cpp_options = list(stan_threads = FALSE), # Enable threading
                                 stanc_options = list("O1")) # Basic optimization

  # Sample from the posterior distribution using MCMC
  fit_bernoulli <- mod_bernoulli$sample(
    data = data_for_stan,          # The data list we prepared
    seed = 123,                    # For reproducible MCMC sampling
    chains = 4,                    # Number of parallel Markov chains (recommend 4)
    parallel_chains = 4, # Run chains in parallel
    #threads_per_chain = 1,         # For within-chain parallelization (usually 1 is fine)
    iter_warmup = 1000,            # Number of warmup iterations per chain (discarded)
    iter_sampling = 2000,          # Number of sampling iterations per chain (kept)
    refresh = 0,                   # How often to print progress updates. 0 means never since the model is so fast
    max_treedepth = 10,            # Controls complexity of MCMC steps (adjust if needed)
    adapt_delta = 0.8              # Target acceptance rate (adjust if divergences occur)
  )

  # Save the fitted model object
  fit_bernoulli$save_object(file = model_file_bernoulli)
  cat("Model fit completed and saved to:", model_file_bernoulli, "\n")
} else {
  # Load existing fitted model object
  fit_bernoulli <- readRDS(model_file_bernoulli)
  cat("Loaded existing model fit from:", model_file_bernoulli, "\n")
}
```

## SUMMARIZE RESULTS & EXTRACT POSTERUIOR DRAWS
```{r}
# Summary
fit_bernoulli$summary()
```

```{r}
# Extract posterior samples and include sampling of the prior:
draws_df_fit_bernoulli <- as_draws_df(fit_bernoulli$draws()) 

# Explicitly extract both parameters
alpha_param <- draws_df_fit_bernoulli$alpha
tau_param <- draws_df_fit_bernoulli$tau
cat("Successfully extracted alpha parameter with", length(alpha_param), "values\n")
cat("Successfully extracted tau parameter with", length(tau_param), "values\n")
```


# MCMC DIAGNOSTICS

```{r}
# 1. Check for Divergent Transitions
diagnostics <- fit_bernoulli$diagnostic_summary()
cat("Divergent transitions per chain:", paste(diagnostics$num_divergent, collapse = ", "), "\n")
```

```{r}
# 2. Check R-hat and ESS metrics - Alpha parameter
fit_summary <- fit_bernoulli$summary(variables = c("alpha", "tau"))
fit_summary <- as.data.frame(fit_summary)

print(fit_summary %>% select(variable, mean, rhat, ess_bulk, ess_tail))
```


```{r}
# 3. Visual Checks (Trace and Rank Plots)
draws_df <- as_draws_df(fit_bernoulli$draws())

# A healthy trace plot is well-mixed with no wandering
p_trace <- mcmc_trace(draws_df, pars = c("alpha", "tau")) + 
  theme_classic() + 
  ggtitle("Trace Plot: Well-mixed chains")

# A healthy rank histogram has uniform bars across all chains
p_rank <- mcmc_rank_hist(draws_df, pars = c("alpha", "tau")) + 
  theme_classic() + 
  ggtitle("Rank Histogram: Chains explored the same space")

p_trace / p_rank

```

# Visualize the posterior distributions for alpha and tau parameters across MCMC iterations
```{r}
ggplot(draws_df_fit_bernoulli, aes(.iteration, alpha, group = .chain, color = .chain)) +
  geom_line() +
  theme_classic()

ggplot(draws_df_fit_bernoulli, aes(.iteration, tau, group = .chain, color = .chain)) +
  geom_line() +
  theme_classic()

```


# PRIOR PREDICTIVE CHECK - Do my priors generate plausible data?
```{r}
# Extract the summary statistic calculated from prior predictive simulations
# 'prior_rep_sum' was calculated within the generated quantities block
draws_gq <- as_draws_df(fit_bernoulli$draws()) # Ensure we have the draws as a dataframe
observed_sum <- sum(data_for_stan$choice) # Total number of 'right' choices in the actual observed data

ppc_prior_ggplot <- ggplot(draws_gq, aes(x = prior_rep_sum)) +
  # Draw the histogram of the model's prior predictions (light red)
  geom_histogram(fill = "#ffbaba", color = "#ff5252", bins = 30, alpha = 0.8) +
  
  # Add the vertical line for the actual observed data (dark red)
  geom_vline(xintercept = observed_sum, color = "darkred", linewidth = 1.2) +
  
  # Labels and theming
  ggtitle("Prior Predictive Check: Total 'Right' Choices") +
  xlab(paste("Total Number of 'Right' Choices (out of", trials, "trials)")) +
  ylab("Frequency") +
  theme_classic()

ppc_prior_ggplot
```
The vertical line (dark red) represents the actual observed data (total ‘right’ choices). If this line falls in a region of the histogram with very low density, it suggests that our priors are not well-aligned with the observed data, which could lead to poor model performance. In this case, the observed data is in the middle of the prior predictive distribution, indicating that our priors are ok and are appropriately centered around plausible values for alpha and tau. This suggests we can continue with our priors as we can assume they reflect our expectations about the parameters based on domain knowledge or previous research.


# PRIOR-POSTERIOR UPDATE CHECK - Did data shift alpha away from prior?
```{r}
# Checking prior alpha to posterior alpha distribution - ideally, the posterior (blue) should be more concentrated around the true value (0.8) than the prior (red), which is more spread out.
ggplot(draws_df) + geom_histogram(aes(x=alpha_prior), colour = "red") + geom_histogram(aes(x = alpha), colour ="blue")

ggplot(draws_df) + geom_histogram(aes(x=tau_prior), colour = "red") + geom_histogram(aes(x = tau), colour ="blue")
```
This is just an alternative way to show the same as the histograms above.
```{r}
# Now let's plot the density for alpha (prior and posterior)
ggplot(draws_df_fit_bernoulli) +
  geom_density(aes(alpha), fill = "blue", alpha = 0.8) +
  geom_density(aes(alpha_prior), fill = "red", alpha = 0.8) +
  geom_vline(xintercept = 0.8, linetype = "dashed", color = "black", linewidth = 1.5) +
  xlab("Alpha") +
  ylab("Posterior Density") +
  theme_classic()
```

```{r}
# plot tau prior and tau true posterior density check
ggplot(draws_df_fit_bernoulli) +
  geom_density(aes(tau), fill = "blue", alpha = 0.8) +
  geom_density(aes(tau_prior), fill = "red", alpha = 0.8) +
  geom_vline(xintercept = 5, linetype = "dashed", color = "black", linewidth = 1.5) +
  xlab("Rate") +
  ylab("Posterior Density") +
  theme_classic()
```

The prior predictive distribution (red histogram) shows the generative space induced by our prior: the wide range of outcomes (total ‘right’ choices) considered possible by the model before seeing the data, driven solely by the prior on alpha_logit and tau_logit.


# POSTERIOR PREDICTIVE CHECKS - Does the Fitted Model Resemble the Data?
```{r}
# Extract posterior predictive simulations from the fitted object
# This matrix has dimensions: [number_of_draws, number_of_trials]
choice_post_rep <- fit_bernoulli$draws("choice_post_rep", format = "matrix")

# Calculate the test statistic T(y) for the *actual* empirical data
observed_sum <- sum(data_for_stan$choice) 

# Calculate the test statistic T(y_rep) for *each* replicated dataset
# We compute the row-wise sum across the posterior draws
post_rep_sums <- rowSums(choice_post_rep)

# Bundle into a tibble for strictly structured plotting
ppc_df <- tibble(
  Replicated_Sum = post_rep_sums
)

# --- Graphical Posterior Predictive Check using ggplot2 ---
ppc_posterior_ggplot <- ggplot(ppc_df, aes(x = Replicated_Sum)) +
  geom_histogram(fill = "skyblue", color = "white", alpha = 0.7, bins =30)+
  geom_vline(xintercept = observed_sum, color = "black", linetype = "dashed", linewidth = 1) +
  ggtitle("Posterior Predictive Check: Total 'Right' Choices") +
  xlab(paste("Total Number of 'Right' Choices (out of", trials, "trials)")) +
  ylab("Frequency (Replicated Datasets)") +
  # Add annotation for the empirical ground truth
  annotate(
    "text", x = observed_sum, y = Inf, 
    label = "Empirical data", 
    vjust = 2, hjust = if(observed_sum > trials/2) 1.1 else -0.1, 
    size = 4, fontface = "bold"
  ) +
  theme_classic()

ppc_posterior_ggplot
```

