---
title: "StanModel"
output: html_document
---

# Setup
```{r}
pacman::p_load(
  tidyverse,    # Data manipulation and plotting
  here,         # Robust file paths
  cmdstanr,     # Stan interface
  rstan,
  posterior,    # Posterior analysis
  bayesplot,    # Plotting MCMC
  furrr,        # Parallel processing
  future,        # Future backend
  patchwork
)
```

```{r}
# Parallel Setup: Leave 2 cores free for system stability
n_cores <- parallel::detectCores() - 2
plan(multisession, workers = n_cores)
```

```{r}
# Optional: Set a default ggplot theme
theme_set(theme_classic())
```

# Simulating Data
```{r}
# Flag to control whether to regenerate simulation/fitting results for the knitted files
# Set to TRUE to rerun everything (takes time!), FALSE to load saved results.
regenerate_simulations <- TRUE # Or TRUE

```

```{r}
# --- Simulate Data for Fitting ---
# Goal: Generate data from known models/parameters to test our fitting procedures.
trials <- 120 # Number of trials per simulated agent
```

# DEFINE AGENTS
```{r}
RandomAgentNoise_f <- function(rate_prob, noise) {
  # Ensure rate is probability (in case log-odds are passed, though not here)
  choice <- rbinom(1, 1, rate_prob) # Base choice
  if (noise > 0 & rbinom(1, 1, noise) == 1) { # Check if noise occurs
    choice <- rbinom(1, 1, 0.5) # Override with random 50/50 choice
  }
  return(choice)
}
```

```{r}
# Define Agent - what kind of player is this simulated data from?
rl_agent_f <- function(opp_choice, estimated_value, alpha =0.4, tau=1) { 
  
  PE = opp_choice - estimated_value;
  estimated_value = estimated_value + (alpha * PE);
    
  # softmax
  p = exp(tau * estimated_value) / (exp(tau * estimated_value) + exp(tau * (1-estimated_value)));

  # choice based on softmax probabilities - prob needs to be a vector as it is for both hands
  choice <- sample(c(0, 1), 1, prob = c(1 - p, p))
  
   return(list(
    choice = choice,
    estimated_value = estimated_value
  ))
}
```

# FULL SIMULATION FUNCTION
```{r} 
full_simulation <- function(alpha, tau, trials =120) {

# Create empty lists for choices and feedback
RA_choices <- rep(NA, trials)
RL_choices <- rep(NA, trials)
feedback_RA <- rep(NA, trials)
feedback_RL <- rep(NA, trials)
estimated_value_tracked <- rep(NA, trials) 
estimated_value_tracked[1] <- 0.5 #tracking "left" hand - estiamted_value[1]
 

RA_choices[1] <- sample(c(0, 1), 1) #choice on trial 1 for agent 1 
RL_choices[1] <- sample(c(0, 1), 1) #choice on trial 1 for agent 2 

for (t in 2:trials) {
  # Feedback for agent 1 (RL) from previous trial (wins if match)
  prev_feedback1 <- ifelse(RL_choices[t - 1] == RA_choices[t - 1], 1, 0)
  feedback_RL[t - 1] <- prev_feedback1

  # Feedback for agent 2 (RA - Opponent) is the opposite (wins if mismatch)
  prev_feedback2 <- 1 - prev_feedback1
  feedback_RA[t - 1] <- prev_feedback2

  # Both agents choose based on their own previous feedback
  RA_choices[t] <- RandomAgentNoise_f(rate_prob=0.8, noise=0.2) #RA agent choice
  
  # RL agent choice and update estimated value based on previous feedback
  result <- rl_agent_f(
    opp_choice = RA_choices[t - 1], # opponents choice to drive the PE in the rl_agent_f
    estimated_value = estimated_value_tracked[t - 1],
    alpha=alpha,
    tau = tau
  )
  
  # RL agent choice and update estimated value defined for current trial
  RL_choices[t] <- result$choice
  estimated_value_tracked[t] <- result$estimated_value
}

# Defining the feedback across trials
feedback_RL[trials] <- ifelse(RL_choices[trials] == RA_choices[trials], 1, 0)
feedback_RA[trials] <- ifelse(RL_choices[trials] == RA_choices[trials], 0, 1)

# Create dataframe
df_RA_RL <- tibble(
  trial = 1:trials,
  Choice_RL = RL_choices,
  Opponent_RA = RA_choices,
  Feedback_RL = feedback_RL, # 1 if agent_RL won
  Feedback_RA = feedback_RA, # 1 if opponent_RA won
  RL_estimated_value_left = estimated_value_tracked,
  RL_estimated_value_right = 1-estimated_value_tracked,
  true_alpha=alpha,
  true_tau= tau
) %>% 
  group_by(true_alpha, true_tau) %>%
  mutate(cumulative_rate = cumsum(Choice_RL) / row_number()) %>%
  ungroup()

return(df_RA_RL)
}
```

# GENERATE DATA ACROSS DIFF PARAMETERS
```{r}
# --- Generate Data Across Different Parameters ---
# Define relative paths for saving/loading simulation data
sim_data_file <- here("Assignment02/data/Simulation_01.csv") 


if (regenerate_simulations || !file.exists(sim_data_file)) {
  cat("Generating new simulation data...\n")
  # Use expand_grid for cleaner parameter combinations
  param_grid <- expand_grid(
    alpha = seq(0, 1, 0.1), # alpha levels
    tau = c(0.5, 1, 5, 10, 15, 20)     # Tau temp levels
  )

  
  # Use map_dfr for efficient simulation across parameters
  d <- pmap_dfr(param_grid, full_simulation) 

  # Save the generated data
  write_csv(d, sim_data_file)
  cat("Generated new simulation data and saved to", sim_data_file, "\n")
} else {
  # Load existing simulation data
  d <- read_csv(sim_data_file)
  cat("Loaded existing simulation data from", sim_data_file, "\n")
}
```

# VISUALISE SIMULATED DATA
```{r}
# --- Visualize Simulated Data ---
# Plot cumulative alpha, faceted by tau level
p1 <- ggplot(d, aes(x = trial, y = cumulative_rate, group = true_tau, color = true_tau)) +
  geom_line(alpha = 0.8) +
  geom_hline(yintercept = 0.5, linetype = "dashed") +
  scale_color_viridis_c() + # Use perceptually uniform color scale
  facet_wrap(~paste("Alpha =", true_alpha)) + # Facet by noise level
  labs(
    title = "Simulated Choice Behavior (Cumulative Rate)",
    subtitle = "Agents with different alpha and tau levels",
    x = "Trial Number",
    y = "Cumulative Proportion Choosing 'Right'",
    color = "True Tau Rate"
  ) +
  theme_minimal() +
  ylim(0, 1)

print(p1)

```

# SELECT ONE ALPHA TAU COMBINATION TO FIT TO STAN MODEL
```{r}
# isolate for one alpha tau combination parameters then fit to the stan model

# Subset data for one specific agent simulation (alpha=0.8, tau=0.5)
d1 <- d %>% filter(true_alpha == 0.8, true_tau == 1) 

# Stan requires data in a list format
data_for_stan <- list(
  players = 1,
  n = trials,     # Number of trials (integer)
  choice = d1$Choice_RL,   # Vector of choices (0s and 1s) for the RL agent
  opp_choice = d1$Opponent_RA # Vector of opponent choices (0s and 1s) for the RA agent
)

str(data_for_stan) # Show the structure of the list

```


# Then need R framework to create the data, fit it to Stan, and then extract the posterior draws and visualise them.

#RUN STAN MODEL feeding it simulated data from above 
```{r}
## --- Compile and Fit ---
# Specify model file path
stan_file_bernoulli <- here("Assignment02/src/Stan_Model.stan")
# Define path for saving fitted model object
stan_results_dir <- "Assignment02/src/Stan_results"
model_file_bernoulli <- here(stan_results_dir, "SimpleBernoulli.rds")

# create folder if it's not there
if (!dir.exists(stan_results_dir)) dir.create(stan_results_dir)

if (regenerate_simulations || !file.exists(model_file_bernoulli)) {
  cat("Compiling and fitting Stan model (Bernoulli)...\n")
  # Compile the Stan model (only needs to be done once unless code changes)
  mod_bernoulli <- cmdstan_model(stan_file_bernoulli,
                                 cpp_options = list(stan_threads = FALSE), # Enable threading
                                 stanc_options = list("O1")) # Basic optimization

  # Sample from the posterior distribution using MCMC
  fit_bernoulli <- mod_bernoulli$sample(
    data = data_for_stan,          # The data list we prepared
    seed = 123,                    # For reproducible MCMC sampling
    chains = 4,                    # Number of parallel Markov chains (recommend 4)
    parallel_chains = 4, # Run chains in parallel
    #threads_per_chain = 1,         # For within-chain parallelization (usually 1 is fine)
    iter_warmup = 1000,            # Number of warmup iterations per chain (discarded)
    iter_sampling = 2000,          # Number of sampling iterations per chain (kept)
    refresh = 0,                   # How often to print progress updates. 0 means never since the model is so fast
    max_treedepth = 10,            # Controls complexity of MCMC steps (adjust if needed)
    adapt_delta = 0.8              # Target acceptance rate (adjust if divergences occur)
  )

  # Save the fitted model object
  fit_bernoulli$save_object(file = model_file_bernoulli)
  cat("Model fit completed and saved to:", model_file_bernoulli, "\n")
} else {
  # Load existing fitted model object
  fit_bernoulli <- readRDS(model_file_bernoulli)
  cat("Loaded existing model fit from:", model_file_bernoulli, "\n")
}
```

## SUMMARIZE RESULTS & EXTRACT POSTERUIOR DRAWS
```{r}
# Summary
fit_bernoulli$summary()
```

```{r}
# Extract posterior samples and include sampling of the prior:
draws_df_fit_bernoulli <- as_draws_df(fit_bernoulli$draws()) 

# Explicitly extract both parameters
alpha_param <- draws_df_fit_bernoulli$alpha
tau_param <- draws_df_fit_bernoulli$tau
cat("Successfully extracted alpha parameter with", length(alpha_param), "values\n")
cat("Successfully extracted tau parameter with", length(tau_param), "values\n")
```


# MCMC DIAGNOSTICS

```{r}
# 1. Check for Divergent Transitions
diagnostics <- fit_bernoulli$diagnostic_summary()
cat("Divergent transitions per chain:", paste(diagnostics$num_divergent, collapse = ", "), "\n")
```

```{r}
# 2. Check R-hat and ESS metrics - Alpha and Tau parameter
fit_summary <- fit_bernoulli$summary(variables = c("alpha", "tau"))
fit_summary <- as.data.frame(fit_summary)

print(fit_summary %>% select(variable, mean, rhat, ess_bulk, ess_tail))
```


```{r}
# VISUAL MODEL CONVERGENCE CHECKS

# 3. Visual Checks (Trace and Rank Plots)
draws_df <- as_draws_df(fit_bernoulli$draws())

# A healthy trace plot is well-mixed with no wandering
p_trace <- mcmc_trace(draws_df, pars = c("alpha", "tau")) + 
  theme_classic() + 
  ggtitle("Trace Plot: Well-mixed chains")

# A healthy rank histogram has uniform bars across all chains - SHOULD BE UNIFORMALLY DISTRIBUTED 
p_rank <- mcmc_rank_hist(draws_df, pars = c("alpha", "tau")) + 
  theme_classic() + 
  ggtitle("Rank Histogram: Chains explored the same space")

p_trace / p_rank

```


# QUALITY ASSESSMENT - PREDICTIVE CHECKS

## PRIOR PREDICTIVE CHECK - Do my priors generate plausible data?
```{r}
# Extract the summary statistic calculated from prior predictive simulations
# 'prior_rep_sum' was calculated within the generated quantities block
draws_gq <- as_draws_df(fit_bernoulli$draws()) # Ensure we have the draws as a dataframe
observed_sum <- sum(data_for_stan$choice) # Total number of 'right' choices in the actual observed data

ppc_prior_ggplot <- ggplot(draws_gq, aes(x = prior_rep_sum)) +
  # Draw the histogram of the model's prior predictions (light red)
  geom_histogram(fill = "#ffbaba", color = "#ff5252", bins = 30, alpha = 0.8) +
  
  # Add the vertical line for the actual observed data (dark red)
  geom_vline(xintercept = observed_sum, color = "darkred", linewidth = 1.2) +
  
  # Labels and theming
  ggtitle("Prior Predictive Check: Total 'Right' Choices") +
  xlab(paste("Total Number of 'Right' Choices (out of", trials, "trials)")) +
  ylab("Frequency") +
  theme_classic()+
  xlim(0,120)

ppc_prior_ggplot
```
The vertical line (dark red) represents the actual observed data (total ‘right’ choices). The pink histogram lines is data generated by our priors. As our priors have not seen the data yet then it is ok that they are not directly on top of the observed data. The data is skewed to pick right by our priors, which is an ok assumption given that the opponent is playing a right biased agent.
Hence, in this case, our priors are ok and are appropriately skewed. This suggests we can continue with our priors as we can assume they reflect our expectations about the parameters based on domain knowledge and the opponent agent.


## PRIOR-POSTERIOR UPDATE CHECK - Did data shift alpha away from prior?
```{r}
# Checking prior alpha to posterior alpha distribution - ideally, the posterior (blue) should be more concentrated around the true value (0.8) than the prior (red), which is more spread out.
ggplot(draws_df) + geom_histogram(aes(x=alpha_prior), fill = "red", colour = "black") + geom_histogram(aes(x = alpha), fill ="blue", colour = "black", alpha = 0.8)

ggplot(draws_df) + geom_histogram(aes(x=tau_prior), fill = "red", colour = "black") + geom_histogram(aes(x = tau), fill ="blue", colour = "black", alpha = 0.8)
```
This is just an alternative way to show the same as the histograms above.
```{r}
# Now let's plot the density for alpha (prior and posterior)
ggplot(draws_df_fit_bernoulli) +
  geom_density(aes(alpha), fill = "blue", alpha = 0.8) +
  geom_density(aes(alpha_prior), fill = "red", alpha = 0.8) +
  geom_vline(xintercept = 0.8, linetype = "dashed", color = "black", linewidth = 1.5) +
  xlab("Alpha") +
  ylab("Posterior Density") +
  theme_classic()
```

```{r}
# plot tau prior and tau true posterior density check
ggplot(draws_df_fit_bernoulli) +
  geom_density(aes(tau), fill = "blue", alpha = 0.8) +
  geom_density(aes(tau_prior), fill = "red", alpha = 0.8) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "black", linewidth = 1.5) +
  xlab("Tau") +
  ylab("Posterior Density") +
  theme_classic()
```

The prior predictive distribution (red histogram) shows the generative space induced by our prior: the wide range of outcomes (total ‘right’ choices) considered possible by the model before seeing the data, driven solely by the prior on alpha_logit and tau_logit.


## POSTERIOR PREDICTIVE CHECKS - Does the Fitted Model Resemble the Data?
```{r}
# Extract posterior predictive simulations from the fitted object
# This matrix has dimensions: [number_of_draws, number_of_trials]
choice_post_rep <- fit_bernoulli$draws("choice_post_rep", format = "matrix")

# Calculate the test statistic T(y) for the *actual* empirical data
observed_sum <- sum(data_for_stan$choice) 

# Calculate the test statistic T(y_rep) for *each* replicated dataset
# We compute the row-wise sum across the posterior draws
post_rep_sums <- rowSums(choice_post_rep)

# Bundle into a tibble for strictly structured plotting
ppc_df <- tibble(
  Replicated_Sum = post_rep_sums
)

# --- Graphical Posterior Predictive Check using ggplot2 ---
ppc_posterior_ggplot <- ggplot(ppc_df, aes(x = Replicated_Sum)) +
  geom_histogram(fill = "skyblue", color = "white", alpha = 0.7, bins =30)+
  geom_vline(xintercept = observed_sum, color = "black", linetype = "dashed", linewidth = 1) +
  ggtitle("Posterior Predictive Check: Total 'Right' Choices") +
  xlab(paste("Total Number of 'Right' Choices (out of", trials, "trials)")) +
  ylab("Frequency (Replicated Datasets)") +
  # Add annotation for the empirical ground truth
  annotate(
    "text", x = observed_sum, y = Inf, 
    label = "Empirical data", 
    vjust = 2, hjust = if(observed_sum > trials/2) 1.1 else -0.1, 
    size = 4, fontface = "bold"
  ) +
  theme_classic()

ppc_posterior_ggplot
```

# PARAMETER RECOVERY
```{r}
recovery_grid <- tidyr::expand_grid(
  n_trials = c(30, 60, 120, 240),
  true_alpha = c(0.1, 0.5, 0.9),
  true_tau = c(1, 5, 15)
)

recovery_file <- file.path("Stan_results/ParameterRecovery.rds")


if (regenerate_simulations || !file.exists(recovery_file)) {
  
  # 2. Define the isolated recovery pipeline
  run_recovery <- function(n_trials, true_alpha, true_tau) {
    # Generate synthetic empirical data
    sim_data <- full_simulation(alpha = true_alpha, tau = true_tau, trials = n_trials)
    
    # Structure the data block for Stan
    rec_data_list <- list(
      players = 1,
      n = n_trials,
      choice = sim_data$Choice_RL[1:n_trials],
      opp_choice = sim_data$Opponent_RA[1:n_trials]
    )
    
    # Fit the generative model
    fit <- mod_bernoulli$sample(
      data = rec_data_list,
      chains = 1,
      iter_warmup = 1000,
      iter_sampling = 2000,
      refresh = 0,
      show_messages = FALSE
    )
    
    # Extract structural draws immediately to avoid temporary file fragility
    draws_df <- fit$draws(c("alpha", "tau"), format = "df")
    
    tibble::tibble(
      trials = n_trials,
      true_alpha = true_alpha,
      alpha_prob = draws_df$alpha,
      true_tau = true_tau,
      tau_prob = draws_df$tau,
    )
  }

  # 3. Establish the parallel backend and execute mapping
  future::plan(future::multisession, workers = 2)
    
  # Replace the parallel section with this temporarily
  recovery_draws <- purrr::pmap(
    recovery_grid,
    run_recovery
  ) |> 
    dplyr::bind_rows()

  future::plan(future::sequential)
  
  saveRDS(recovery_draws, recovery_file)
  
} else {
  recovery_draws <- readRDS(recovery_file)
  cat("Loaded existing 2D recovery draws from:", recovery_file, "\n")
}
```

```{r}
# recovert summary
recovery_summary <- recovery_draws |> 
  dplyr::group_by(trials, true_alpha, true_tau) |> 
  dplyr::summarize(
    post_mean_alpha = mean(alpha_prob),
    post_mean_tau = mean(tau_prob),
    lower_95_alpha = quantile(alpha_prob, 0.025),
    upper_95_alpha = quantile(alpha_prob, 0.975),
    lower_95_tau = quantile(tau_prob, 0.025),
    upper_95_tau = quantile(tau_prob, 0.975),
    .groups = "drop"
  ) |> 
  dplyr::mutate(
    facet_label_alpha = paste0("True α = ", true_alpha),  # ← add this
    facet_label_tau   = paste0("True τ = ", true_tau)     # ← add this
  )
```

```{r}
# Alpha recovery - facet by trials, colour by true_tau
p_alpha_facet <- ggplot(recovery_summary, 
                        aes(x = true_alpha, y = post_mean_alpha, color = factor(true_tau))) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black") +
  geom_pointrange(aes(ymin = lower_95_alpha, ymax = upper_95_alpha),
                  linewidth = 0.8, size = 0.6,
                  position = position_dodge(width = 0.02)) +
  facet_wrap(~ paste("N Trials =", trials)) +
  scale_color_viridis_d(option = "mako", begin = 0.2, end = 0.8) +
  labs(
    title = expression("Parameter Recovery: Learning Rate " * alpha),
    subtitle = expression("Faceted by trial count, coloured by true " * tau),
    x = expression("True " * alpha),
    y = expression("Recovered " * alpha * " (Mean & 95% CrI)"),
    color = expression("True " * tau)
  ) +
  coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
  theme_classic() +
  theme(
    strip.background = element_rect(fill = "gray95", color = NA),
    strip.text = element_text(face = "bold", size = 11),
    panel.grid.major = element_line(color = "gray90", linetype = "dotted")
  )

# Tau recovery - facet by trials, colour by true_alpha
p_tau_facet <- ggplot(recovery_summary, 
                      aes(x = true_tau, y = post_mean_tau, color = factor(true_alpha))) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "black") +
  geom_pointrange(aes(ymin = lower_95_tau, ymax = upper_95_tau),
                  linewidth = 0.8, size = 0.6,
                  position = position_dodge(width = 0.3)) +
  facet_wrap(~ paste("N Trials =", trials)) +
  scale_color_viridis_d(option = "rocket", begin = 0.2, end = 0.8) +
  labs(
    title = expression("Parameter Recovery: Inverse Temperature " * tau),
    subtitle = expression("Faceted by trial count, coloured by true " * alpha),
    x = expression("True " * tau),
    y = expression("Recovered " * tau * " (Mean & 95% CrI)"),
    color = expression("True " * alpha)
  ) +
  coord_cartesian(xlim = c(0, 20), ylim = c(0, 20)) +
  theme_classic() +
  theme(
    strip.background = element_rect(fill = "gray95", color = NA),
    strip.text = element_text(face = "bold", size = 11),
    panel.grid.major = element_line(color = "gray90", linetype = "dotted")
  )

print(p_alpha_facet)
print(p_tau_facet)
```
The Tau plot indicates that when alpha is low (0.1) the recovery of tau is good as it falls on the line and has the smallest CIs. Performance also improves for recovery of low tau values as the number of trials increases. A low learning rate may mean that the data is more deterministic and thus easier to recover tau, whereas when alpha is high (0.9) the data is more stochastic and thus harder to recover tau.
The alpha plot indicates that when tau is high (15) the recovery of alpha is good as it falls on the line and has the smallest CIs. Increasing the number of trials does not massively influence parameter recovery performance. The deterministic tau may mean that it is easier to recover alpha as the data is less noisy, whereas when tau is low (1) the data is more stochastic and thus harder to recover alpha.

# SIMULATION BASED CALIBRATION (SBC)
"Is my model self-consistent with its prior?"

This tests whether your entire Bayesian workflow is coherent — if you repeatedly draw from the prior, simulate data, and refit, the posterior should "contain" the true value uniformly. 
```{r}
# LOAD PACKAGES
#pacman::p_load(devtools)
remotes::install_github("hyunjimoon/SBC")
```

```{r}
library(SBC)

sbc_results_file <- file.path("Stan_results/sbc_results.rds")

if (regenerate_simulations || !file.exists(sbc_results_file)) {
  
# 1. Define the exact mathematical Generative Function
  sbc_generator_fn <- function() {
    # Draw from the structural prior (on the unconstrained space)
    alpha_logit_true <- rnorm(1, mean = 0, sd = 1.5)
    true_alpha <- plogis(alpha_logit_true)
    tau_logit_true <- rnorm(1, mean = 0, sd = 1.5)
    true_tau <- plogis(tau_logit_true) * 20 # Scale to
    
    # Simulate empirical reality (N = 120 trials)
    n_trials <- 120
    sim_data <- full_simulation(alpha = true_alpha, tau = true_tau, trials = n_trials)
    
    # Format strictly required by the SBC package
    list(
      variables = list(alpha_logit = alpha_logit_true, alpha = true_alpha, tau_logit = tau_logit_true, tau = true_tau),
      generated = list(
        players = 1,
        n = n_trials,
        choice = sim_data$Choice_RL[1:n_trials],
        opp_choice = sim_data$Opponent_RA[1:n_trials]
      )
    )
  }

  # Cast the raw function into a formal SBC generator object
  sbc_generator_obj <- SBC::SBC_generator_function(sbc_generator_fn)

  # 2. Register the CmdStanR backend for SBC
  sbc_backend <- SBC::SBC_backend_cmdstan_sample(
    mod_bernoulli, 
    chains = 2,           
    iter_warmup = 500, 
    iter_sampling = 1000,
    refresh = 0,
    show_messages = FALSE
  )

  # 3. Generate datasets and compute rank statistics using parallelization
  future::plan(future::multisession, workers = 2)
  
  # Generate 1000 synthetic realities using the formally cast object!
  sbc_datasets <- SBC::generate_datasets(sbc_generator_obj, n_sims = 200)
  
  # Fit the model 1000 times and extract rank statistics
  sbc_results <- SBC::compute_SBC(sbc_datasets, sbc_backend)
  
  future::plan(future::sequential)
  
  saveRDS(sbc_results, sbc_results_file)
  cat("SBC completed. Results saved to:", sbc_results_file, "\n")
  
} else {
  sbc_results <- readRDS(sbc_results_file)
  cat("Loaded existing SBC results from:", sbc_results_file, "\n")
}
```

```{r}
# 4. Visualize the Calibration Geometry

# Plot A: Rank Histogram

p_sbc_hist <- SBC::plot_rank_hist(sbc_results, variables = c("alpha", "tau")) + 
  ggtitle("SBC Rank Histogram: Algorithmic Calibration") +
  theme_classic() +
  theme(strip.background = element_rect(fill = "gray95", color = "NA"))

# Plot B: Empirical Cumulative Distribution Function (ECDF)

p_sbc_ecdf <- SBC::plot_ecdf(sbc_results, variables = c("alpha", "tau"))  +
  ggtitle("SBC ECDF: Deviation from Uniformity") +
  theme_classic() +
  theme(strip.background = element_rect(fill = "gray95", color = "NA"))

# Combine plots  using patchwork
p_sbc_combined <- p_sbc_hist / p_sbc_ecdf
p_sbc_combined
```
Interpretation of the SBC Geometric Diagnostics: We are interrogating the algorithm for structural bias.

Rank Histogram (Top): The gray shaded region represents the 95% binomial confidence bounds for a perfectly uniform distribution.
If the bars form a U-shape, our model is OVERCONFIDENT. The posterior is pathologically narrow, and the true value keeps falling in the extreme tails.
If the bars form an Inverse U-shape (a hump), our model is UNDERCONFIDENT. The posterior is too wide.
If the bars skew left or right, our model is BIASED, systematically under- or over-estimating the parameter.

ECDF Plot (Bottom): The blue line represents our empirical CDF, and the gray ellipse is the acceptable uniformity bound. If the blue line breaks outside the ellipse, the HMC sampler has failed to capture the true geometry of the prior space.


# POSTERIOR SBC

"Is my model consistent with the data I actually observed?"

This tests whether your model is a plausible generative process for your actual observed data — i.e., data simulated from your posterior should look statistically indistinguishable from what you observed, and refitting on it should recover those parameters.

If PSBC passes but Prior SBC fails, it means your model is only reliable for data similar to what you observed. If PSBC fails, you cannot trust the specific estimates for your current experiment, even if the model works “in general”.

Here:
Sampling from posterior draws with replacement
Extracting true_alpha / true_tau from each sampled row to condition the simulation
Passing the simulated data back as a Stan-compatible list
```{r}
# 1. Extract posterior draws
post_draws_raw <- posterior::as_draws_matrix(
  fit_bernoulli$draws(c("alpha", "tau"))
)

# 2. PSBC generator
generate_psbc_data <- function(n_sims = 100) {
  
  indices <- sample(1:nrow(post_draws_raw), n_sims, replace = TRUE)
  sampled_vars <- post_draws_raw[indices, , drop = FALSE]
  
  gen_list <- vector("list", n_sims)
  
  for (i in 1:n_sims) {
    
    # Draw from the posterior "true" values
    true_alpha <- as.numeric(sampled_vars[i, "alpha"])
    true_tau   <- as.numeric(sampled_vars[i, "tau"])
    
    # Generate synthetic dataset
    sim_data <- full_simulation(
      alpha = true_alpha,
      tau = true_tau,
      trials = data_for_stan$n
    )
    
    # --- Format for Stan ---
    gen_list[[i]] <- list(
      players = 1,
      n = data_for_stan$n,
      choice = sim_data$Choice_RL,
      opp_choice = sim_data$Opponent_RA,
      prior_alpha_m = 0,
      prior_alpha_sd = 1.5,
      prior_tau_m = 0,
      prior_tau_sd = 1.5
    )
  }
  
  SBC::SBC_datasets(
    variables = sampled_vars,
    generated = gen_list
  )
}

psbc_results_file <- file.path("Stan_results/psbc_results.rds")


if (regenerate_simulations || !file.exists(psbc_results_file)) {
  psbc_datasets <- generate_psbc_data(n_sims = 100)
  future::plan(future::sequential)
  psbc_results <- SBC::compute_SBC(psbc_datasets, sbc_backend)
  saveRDS(psbc_results, psbc_results_file)
} else {
  psbc_results <- readRDS(psbc_results_file)
  cat("Loaded existing PSBC results from:", psbc_results_file, "\n")
}

```

```{r}
# Visualize Rank Histograms for local calibration
p_psbc_hist <- SBC::plot_rank_hist(psbc_results, variables = c("alpha", "tau")) +
  ggtitle("Posterior SBC: Reinforcement Learning Agent") +
  theme_classic()

p_psbc_ecdf <- SBC::plot_ecdf(psbc_results, variables = c("alpha", "tau"))  +
  ggtitle("Posterior SBC ECDF: Deviation from Uniformity") +
  theme_classic() +
  theme(strip.background = element_rect(fill = "gray95", color = "NA"))

# Combine plots  using patchwork
p_psbc_combined <- p_psbc_hist / p_psbc_ecdf
p_psbc_combined
```