---
title: "SimulationScript"
output: html_document
---

```{r}
pacman::p_load(
  tidyverse,    # Data manipulation and plotting
  here,         # Robust file paths
  cmdstanr,     # Stan interface
  rstan,
  posterior,    # Posterior analysis
  bayesplot,    # Plotting MCMC
  furrr,        # Parallel processing
  future,        # Future backend
  patchwork
)

main_path <- "/Users/magnusseverin/Documents/GitHub/Advanced_Cognitive_Modeling/Assignment04/"

setwd(main_path)

```


```{r}
# Parallel Setup: Leave 2 cores free for system stability
n_cores <- parallel::detectCores() - 2
plan(multisession, workers = n_cores)
```

```{r}
# Optional: Set a default ggplot theme
theme_set(theme_classic())
```

# Simulating Data
```{r}
# Flag to control whether to regenerate simulation/fitting results for the knitted files
# Set to TRUE to rerun everything (takes time!), FALSE to load saved results.
regenerate_simulations <- TRUE 

library(tidyverse)
set.seed(42)
```

Responses can be 1-4
Response 1 = Not dangerous and not nutritious
Response 2 = Not dangerous and nutritious
Response 3 = Dangerous and not nutritious
Response 4 = Dangerous and nutritious

BUT the associations between feature and category varies over session...

In the first session: danger depends on the alien having spots AND eyes on stalks (feature 1 AND feature 3 both being 1)

In the second session: danger depends on arms being up (feature 4 being 1)

In the third session: danger depends on arms up and green color (feature 4 and 5 being both 1)

```{r}
# Generate all 32 combinations of 5 binary features
features <- expand.grid(
  feature1 = 0:1,  # spots
  feature2 = 0:1,  # slim legs
  feature3 = 0:1,  # eyes on stalks
  feature4 = 0:1,  # arms up
  feature5 = 0:1   # green color
)

# Replicate for 3 sessions 
sim_grid <- bind_rows(
  mutate(features, session = 1),
  mutate(features, session = 2),
  mutate(features, session = 3)
) %>%
  mutate(
    danger = case_when(
      session == 1 ~ as.integer(feature1 == 1 & feature3 == 1),  # spots AND eyes on stalks
      session == 2 ~ as.integer(feature4 == 1),                   # arms up
      session == 3 ~ as.integer(feature4 == 1 & feature5 == 1)    # arms up AND green
    )
  ) %>%
  select(session, feature1, feature2, feature3, feature4, feature5, danger)
```

```{r}
sim_grid <- sim_grid |> filter(session == 2)
```


# Experimental Data 
```{r}
exp_data <- read_csv(glue::glue("{main_path}data/AlienData.txt")) 
```
