---
title: "SimulationScript"
output: html_document
---

```{r}
pacman::p_load(
  tidyverse,    # Data manipulation and plotting
  here,         # Robust file paths
  cmdstanr,     # Stan interface
  rstan,
  posterior,    # Posterior analysis
  bayesplot,    # Plotting MCMC
  furrr,        # Parallel processing
  future,        # Future backend
  patchwork
)

main_path <- "/Users/mattie/Desktop/Masters/ACM/AdvancedCognitiveModelling/Advanced_Cognitive_Modeling/Assignment04/"

knitr::opts_knit$set(root.dir = main_path)

```


```{r}
# Parallel Setup: Leave 2 cores free for system stability
n_cores <- parallel::detectCores() - 2
plan(multisession, workers = n_cores)
```

```{r}
# Optional: Set a default ggplot theme
theme_set(theme_classic())
```

# Simulating Data
```{r}
# Flag to control whether to regenerate simulation/fitting results for the knitted files
# Set to TRUE to rerun everything (takes time!), FALSE to load saved results.
regenerate_simulations <- TRUE 

library(tidyverse)
set.seed(42)
```

Responses can be 1-4
Response 1 = Not dangerous and not nutritious
Response 2 = Not dangerous and nutritious
Response 3 = Dangerous and not nutritious
Response 4 = Dangerous and nutritious

BUT the associations between feature and category varies over session...

In the first session: danger depends on the alien having spots AND eyes on stalks (feature 1 AND feature 3 both being 1)

In the second session: danger depends on arms being up (feature 4 being 1)

In the third session: danger depends on arms up and green color (feature 4 and 5 being both 1)

```{r}
# Generate all 32 combinations of 5 binary features
features <- expand.grid(
  feature1 = 0:1,  # spots
  feature2 = 0:1,  # slim legs
  feature3 = 0:1,  # eyes on stalks
  feature4 = 0:1,  # arms up
  feature5 = 0:1   # green color
)

# Replicate for 3 sessions 
sim_grid <- bind_rows(
  mutate(features, session = 1),
  mutate(features, session = 2),
  mutate(features, session = 3)
) %>%
  mutate(
    danger = case_when(
      session == 1 ~ as.integer(feature1 == 1 & feature3 == 1),  # spots AND eyes on stalks
      session == 2 ~ as.integer(feature4 == 1),                   # arms up
      session == 3 ~ as.integer(feature4 == 1 & feature5 == 1)    # arms up AND green
    )
  ) %>%
  select(session, feature1, feature2, feature3, feature4, feature5, danger)
```

```{r}
sim_grid <- sim_grid |> filter(session == 2)
# mutate an stimulus id column and place first in list
sim_grid <- sim_grid %>% 
  mutate(ID = row_number()) %>% 
  select(ID, everything())
```

SIMILARITY FUNCITON
```{r}
# Exponential similarity kernel
similarity <- function(dist_val, c) {
  exp(-c * dist_val)
}
```

DISTANCE FUNCTION
```{r}
# City-block distance (r = 1), attention-weighted
distance <- function(vect1, vect2, w) {
  sum(w * abs(vect1 - vect2))
}
```


GENERATIVE GCM MODEL FUNCTION
```{r}
# Generative GCM model function.
gcm_simulate <- function(w,            # attention weights (sums to 1)
                         c,            # sensitivity (natural scale, strictly positive) based on distance
                         bias,         # response bias towards category 1
                         features,      # matrix of stimulus features (trials x features)
                         danger,       # true category labels (0 or 1)
                         quiet = TRUE) {
  
  obs       <- as.matrix(features)
  ntrials   <- nrow(obs)
  nfeatures <- ncol(obs)


  # create empty vectors (and matrices) to store results  
  prob_dan1    <- numeric(ntrials) # will be probability of responding "dangerous" on each trial
  sim_response <- numeric(ntrials) # will be simulated binary responses (0 or 1) based on prob_dan1

  memory_obs <- matrix(NA_real_, nrow = 0, ncol = nfeatures) # Memory of observed stimuli (starts empty)
  memory_dan <- numeric(0) # Memory of observed categories (starts empty)

  # Loop through each trial, simulating responses based on the GCM model
  for (i in seq_len(ntrials)) {
    if (!quiet && i %% 20 == 0) cat("Simulating trial:", i, "\n")

    current_stim <- as.numeric(obs[i, ])
    n_mem        <- nrow(memory_obs)
    has_dan0     <- any(memory_dan == 0)
    has_dan1     <- any(memory_dan == 1)

    # If no memory or only one category in memory, rely on bias. Otherwise, compute similarity-based probs.
    if (n_mem == 0 || !has_dan0 || !has_dan1) {
      prob_dan1[i] <- bias
    } else {
      sims <- vapply(seq_len(n_mem), function(e) {
        similarity(distance(current_stim, memory_obs[e, ], w), c)
      }, numeric(1))

      # Mean similarity per category — removes category-frequency bias.
      m1  <- mean(sims[memory_dan == 1])
      m0  <- mean(sims[memory_dan == 0])
      num <- bias * m1
      den <- num + (1 - bias) * m0

      prob_dan1[i] <- if (den > 1e-9) num / den else bias # avoid division by 0
      prob_dan1[i] <- pmax(1e-9, pmin(1 - 1e-9, prob_dan1[i])) # Ensure probabilities are in (0,1)

    }

    # Simulate a binary response based on the computed probability of category 1 (dangerous)
    sim_response[i] <- rbinom(1, 1, prob_dan1[i])

    # Update memory with the current stimulus and its category label
    memory_obs <- rbind(memory_obs, current_stim)
    memory_dan <- c(memory_dan, danger[i])
  }

  # Return a tibble with the prob of a "dangerous" response & the simulated responses for each trial
  tibble(prob_dan1 = prob_dan1, sim_response = sim_response)
}
```

We are using 3 blocks for each session and subject. The stimuli are shuffled for each block.

```{r}
# Experiment design parameters
n_blocks <- 3
n_stim_per_block <- nrow(features)
total_trials <- n_blocks * n_stim_per_block
```


PER SUBJECT SCHEDULE GENERATION
```{r}
# per subject schedule generation
make_subject_schedule <- function(sim_grid, n_blocks, seed) {
  set.seed(seed)
  n_stim <- nrow(sim_grid) # number of unique stimuli (32)

  # Create a random sequence of stimuli for each block, ensuring all stimuli are presented in each block
  sequence <- unlist(lapply(seq_len(n_blocks), function(b) {
    sample(sim_grid$ID, n_stim, replace = FALSE)
  }))

  tibble(
    trial_within_subject = seq_along(sequence), # trial number for the subject
    block                = rep(seq_len(n_blocks), each = n_stim), # block number
    stimulus_id          = sequence # the stimulus presented on each trial (randomized within blocks)
  ) |>
    left_join(sim_grid, by = c("stimulus_id" = "ID")) |> # add stimulus features and category feedback
    rename(danger_feedback = danger) |> # rename the true category label for clarity
    dplyr::select(trial_within_subject, block, stimulus_id, # select and order relevant columns
                  feature1, feature2, feature3, feature4, feature5, danger_feedback)
}
```

SIMULATING BEHAVIOUR WITH GCM
```{r}
# Simulate behavior with GCM

simulate_gcm_agent <- function(agent_id, w_true, c_true, bias_true,
                               sim_grid, n_blocks, subject_seed) {
  
  schedule <- make_subject_schedule(sim_grid, n_blocks, seed = subject_seed) # generate the stimulus schedule for this agent
  obs      <- schedule |> dplyr::select(feature1, feature2, feature3, feature4, feature5) # extract the stimulus features for each trial
  feedback <- schedule$danger_feedback # extract the true danger labels for each trial

  # apply GCM simulation to generate responses based on the true parameters and the stimulus schedule
  sim_results <- gcm_simulate(
    w = w_true, 
    c = c_true, 
    bias = bias_true,
    features = obs,
    danger = feedback, 
    quiet = TRUE
  )

  schedule |>
    mutate(
      agent_id     = agent_id,
      w1_true      = w_true[1],
      w2_true      = w_true[2],
      c_true       = c_true,
      log_c_true   = log(c_true),
      bias_true    = bias_true,
      prob_dan1    = sim_results$prob_dan1,
      sim_response = sim_results$sim_response,
      correct      = as.integer(danger_feedback == sim_response)
    ) |>
    group_by(agent_id) |>
    mutate(performance = cumsum(correct) / row_number()) |>
    ungroup()
}
```

TESTING DIFFERENT PARAMETER SETTINGS
```{r}
param_grid <- expand_grid(
  w_setting    = list(
        c(0.2, 0.2, 0.2, 0.2, 0.2),        # equal attention to all features
        c(0.4, 0.1, 0.4, 0.05, 0.05),      # attention on features 1 & 3 (session 1 rule)
        c(0.05, 0.05, 0.05, 0.8, 0.05),    # attention on feature 4 (session 2 rule)
        c(0.05, 0.05, 0.05, 0.45, 0.4)     # attention on features 4 & 5 (session 3 rule)
        ),
  c_setting    = c(0.5, 1.0, 2.0, 4.0, 5.0, 15.0),
  bias_setting = 0.5
) |>
  mutate( # Add labels for each parameter setting for later use in plotting
    setting_id = row_number(),
    w_label    = map_chr(w_setting, ~paste(round(.x, 1), collapse = ",")),
    c_label    = paste0("c≈", round(c_setting, 1)),
    label      = paste0("w=(", w_label, "), ", c_label)
  )

n_agents_per_setting <- 5

# Create a simulation plan that expands the parameter grid to include multiple agents per setting, assigning a unique agent ID to each simulated agent.
simulation_plan <- param_grid |>
  uncount(n_agents_per_setting) |>
  mutate(agent_id = row_number())

# create simdata folder
if (!dir.exists(glue::glue("{main_path}/simdata"))) {
  dir.create(glue::glue("{main_path}/simdata"))
}
sim_file <- glue::glue("{main_path}/simdata/simulated_responses.csv")

if (regenerate_simulations || !file.exists(sim_file)) {

  # run simulations in parallel across all parameter settings and agents
  simulated_responses <- future_pmap_dfr(
    list(
      agent_id  = simulation_plan$agent_id,
      w_true    = simulation_plan$w_setting,
      c_true    = simulation_plan$c_setting,
      bias_true = simulation_plan$bias_setting
    ),
    simulate_gcm_agent,
    sim_grid = sim_grid,
    n_blocks      = n_blocks,
    subject_seed  = simulation_plan$agent_id,
    .options = furrr_options(seed = TRUE, packages = c("tidyverse", "dplyr"))
  )

  # Save responses for future use
  plan(sequential)
  write_csv(simulated_responses, sim_file)
  cat("Simulation complete. Results saved.\n")
} else {
  simulated_responses <- read_csv(sim_file, show_col_types = FALSE)
  cat("Loaded existing simulation results.\n")
}
```


# Experimental Data 
```{r}
exp_data <- read_csv(glue::glue("{main_path}data/AlienData.txt")) 
```
