---
title: "SimulationScript"
output: html_document
---

```{r}
pacman::p_load(
  tidyverse,    # Data manipulation and plotting
  here,         # Robust file paths
  cmdstanr,     # Stan interface
  rstan,
  posterior,    # Posterior analysis
  bayesplot,    # Plotting MCMC
  forcats,
  furrr,        # Parallel processing
  future,        # Future backend
  patchwork,
  hexbin
)

main_path <- "/Users/mattie/Desktop/Masters/ACM/AdvancedCognitiveModelling/Advanced_Cognitive_Modeling/Assignment04/"

knitr::opts_knit$set(root.dir = main_path)

```


```{r}
# Parallel Setup: Leave 2 cores free for system stability
n_cores <- parallel::detectCores() - 2
plan(multisession, workers = n_cores)
```

```{r}
# Optional: Set a default ggplot theme
theme_set(theme_classic())
```

# Simulating Data
```{r}
# Flag to control whether to regenerate simulation/fitting results for the knitted files
# Set to TRUE to rerun everything (takes time!), FALSE to load saved results.
regenerate_simulations <- TRUE 

library(tidyverse)
set.seed(42)
```

Responses can be 1-4
Response 1 = Not dangerous and not nutritious
Response 2 = Not dangerous and nutritious
Response 3 = Dangerous and not nutritious
Response 4 = Dangerous and nutritious

BUT the associations between feature and category varies over session...

In the first session: danger depends on the alien having spots AND eyes on stalks (feature 1 AND feature 3 both being 1)

In the second session: danger depends on arms being up (feature 4 being 1)

In the third session: danger depends on arms up and green color (feature 4 and 5 being both 1)

FEATURE GENERATION
```{r}
# Generate all 32 combinations of 5 binary features
features <- expand.grid(
  feature1 = 0:1,  # spots
  feature2 = 0:1,  # slim legs
  feature3 = 0:1,  # eyes on stalks
  feature4 = 0:1,  # arms up
  feature5 = 0:1   # green color
)

# Replicate for 3 sessions 
sim_grid <- bind_rows(
  mutate(features, session = 1),
  mutate(features, session = 2),
  mutate(features, session = 3)
) %>%
  mutate(
    danger = case_when(
      session == 1 ~ as.integer(feature1 == 1 & feature3 == 1),  # spots AND eyes on stalks
      session == 2 ~ as.integer(feature4 == 1),                   # arms up
      session == 3 ~ as.integer(feature4 == 1 & feature5 == 1)    # arms up AND green
    )
  ) %>%
  select(session, feature1, feature2, feature3, feature4, feature5, danger)
```

```{r}
sim_grid <- sim_grid |> filter(session == 2)
# mutate an stimulus id column and place first in list
sim_grid <- sim_grid %>% 
  mutate(ID = row_number()) %>% 
  select(ID, everything())
```

SIMILARITY FUNCTION
```{r}
# Exponential similarity kernel
similarity <- function(dist_val, c) {
  exp(-c * dist_val)
}
```

DISTANCE FUNCTION
```{r}
# City-block distance (r = 1), attention-weighted
distance <- function(vect1, vect2, w) {
  sum(w * abs(vect1 - vect2))
}
```


GENERATIVE GCM MODEL FUNCTION
```{r}
# Generative GCM model function.
gcm_simulate <- function(w,            # attention weights (sums to 1)
                         c,            # sensitivity (natural scale, strictly positive) based on distance
                         bias,         # response bias towards category 1
                         features,     # matrix of stimulus features (trials x features)
                         danger,       # true category labels (0 or 1)
                         quiet = TRUE) {
  
  obs       <- as.matrix(features)
  ntrials   <- nrow(obs)
  nfeatures <- ncol(obs)


  # create empty vectors (and matrices) to store results  
  prob_dan1    <- numeric(ntrials) # will be probability of responding "dangerous" on each trial
  sim_response <- numeric(ntrials) # will be simulated binary responses (0 or 1) based on prob_dan1

  memory_obs <- matrix(NA_real_, nrow = 0, ncol = nfeatures) # Memory of observed stimuli (starts empty)
  memory_dan <- numeric(0) # Memory of observed categories (starts empty)

  # Loop through each trial, simulating responses based on the GCM model
  for (i in seq_len(ntrials)) {
    if (!quiet && i %% 20 == 0) cat("Simulating trial:", i, "\n")

    current_stim <- as.numeric(obs[i, ])
    n_mem        <- nrow(memory_obs)
    has_dan0     <- any(memory_dan == 0)
    has_dan1     <- any(memory_dan == 1)

    # If no memory or only one category in memory, rely on bias. Otherwise, compute similarity-based probs.
    if (n_mem == 0 || !has_dan0 || !has_dan1) {
      prob_dan1[i] <- bias
    } else {
      sims <- vapply(seq_len(n_mem), function(e) {
        similarity(distance(current_stim, memory_obs[e, ], w), c)
      }, numeric(1))

      # Mean similarity per category — removes category-frequency bias.
      m1  <- mean(sims[memory_dan == 1])
      m0  <- mean(sims[memory_dan == 0])
      num <- bias * m1
      den <- num + (1 - bias) * m0

      prob_dan1[i] <- if (den > 1e-9) num / den else bias # avoid division by 0
      prob_dan1[i] <- pmax(1e-9, pmin(1 - 1e-9, prob_dan1[i])) # Ensure probabilities are in (0,1)

    }

    # Simulate a binary response based on the computed probability of category 1 (dangerous)
    sim_response[i] <- rbinom(1, 1, prob_dan1[i])

    # Update memory with the current stimulus and its category label
    memory_obs <- rbind(memory_obs, current_stim)
    memory_dan <- c(memory_dan, danger[i])
  }

  # Return a tibble with the prob of a "dangerous" response & the simulated responses for each trial
  tibble(prob_dan1 = prob_dan1, sim_response = sim_response)
}
```

We are using 3 blocks for each session and subject. The stimuli are shuffled for each block.

```{r}
# Experiment design parameters
n_blocks <- 3
n_stim_per_block <- nrow(features)
total_trials <- n_blocks * n_stim_per_block
```


PER SUBJECT SCHEDULE GENERATION
```{r}
# per subject schedule generation
make_subject_schedule <- function(sim_grid, n_blocks, seed) {
  set.seed(seed)
  n_stim <- nrow(sim_grid) # number of unique stimuli (32)

  # Create a random sequence of stimuli for each block, ensuring all stimuli are presented in each block
  sequence <- unlist(lapply(seq_len(n_blocks), function(b) {
    sample(sim_grid$ID, n_stim, replace = FALSE)
  }))

  tibble(
    trial_within_subject = seq_along(sequence), # trial number for the subject
    block                = rep(seq_len(n_blocks), each = n_stim), # block number
    stimulus_id          = sequence # the stimulus presented on each trial (randomized within blocks)
  ) |>
    left_join(sim_grid, by = c("stimulus_id" = "ID")) |> # add stimulus features and category feedback
    rename(danger_feedback = danger) |> # rename the true category label for clarity
    dplyr::select(trial_within_subject, block, stimulus_id, # select and order relevant columns
                  feature1, feature2, feature3, feature4, feature5, danger_feedback)
}
```

SIMULATING BEHAVIOUR WITH GCM
```{r}
# Simulate behavior with GCM

simulate_gcm_agent <- function(agent_id, w_true, c_true, bias_true,
                               sim_grid, n_blocks, subject_seed) {
  
  schedule <- make_subject_schedule(sim_grid, n_blocks, seed = subject_seed) # generate the stimulus schedule for this agent
  obs      <- schedule |> dplyr::select(feature1, feature2, feature3, feature4, feature5) # extract the stimulus features for each trial
  feedback <- schedule$danger_feedback # extract the true danger labels for each trial

  # apply GCM simulation to generate responses based on the true parameters and the stimulus schedule
  sim_results <- gcm_simulate(
    w = w_true, 
    c = c_true, 
    bias = bias_true,
    features = obs,
    danger = feedback, 
    quiet = TRUE
  )

  schedule |>
    mutate(
      agent_id     = agent_id,
      w4_true      = w_true[4],
      c_true       = c_true,
      log_c_true   = log(c_true),
      bias_true    = bias_true,
      prob_dan1    = sim_results$prob_dan1,
      sim_response = sim_results$sim_response,
      correct      = as.integer(danger_feedback == sim_response)
    ) |>
    group_by(agent_id) |>
    mutate(performance = cumsum(correct) / row_number()) |>
    ungroup()
}
```

TESTING DIFFERENT PARAMETER SETTINGS
```{r}
param_grid <- expand_grid(
  w_setting    = list(
        c(0.2, 0.2, 0.2, 0.2, 0.2), # equal attention to all features
        c(0.0, 0.0, 0.0, 1.0, 0.0), # extreme attention on feature 4 (session 2 rule)
        c(0.1, 0.1, 0.1, 0.6, 0.1), # middle attention on feature 4 
        c(0.05, 0.05, 0.05, 0.8, 0.05) # strong attention on feature 4 
        ),
  c_setting    = c(0.5, 1.0, 2.0, 4.0, 5.0, 15.0),
  bias_setting = 0.5
) |>
  mutate( # Add labels for each parameter setting for later use in plotting
    setting_id = row_number(),
    w_label    = map_chr(w_setting, ~paste(round(.x, 1), collapse = ",")),
    c_label    = paste0("c≈", round(c_setting, 1)),
    label      = paste0("w=(", w_label, "), ", c_label)
  )

n_agents_per_setting <- 5

# Create a simulation plan that expands the parameter grid to include multiple agents per setting, assigning a unique agent ID to each simulated agent.
simulation_plan <- param_grid |>
  uncount(n_agents_per_setting) |>
  mutate(agent_id = row_number())

# create simdata folder
if (!dir.exists(glue::glue("{main_path}/simdata"))) {
  dir.create(glue::glue("{main_path}/simdata"))
}
sim_file <- glue::glue("{main_path}/simdata/simulated_responses.csv")

if (regenerate_simulations || !file.exists(sim_file)) {

  # run simulations in parallel across all parameter settings and agents
  simulated_responses <- future_pmap_dfr(
    list(
      agent_id  = simulation_plan$agent_id,
      w_true    = simulation_plan$w_setting,
      c_true    = simulation_plan$c_setting,
      bias_true = simulation_plan$bias_setting
    ),
    simulate_gcm_agent,
    sim_grid = sim_grid,
    n_blocks      = n_blocks,
    subject_seed  = simulation_plan$agent_id,
    .options = furrr_options(seed = TRUE, packages = c("tidyverse", "dplyr"))
  )

  # Save responses for future use
  plan(sequential)
  write_csv(simulated_responses, sim_file)
  cat("Simulation complete. Results saved.\n")
} else {
  simulated_responses <- read_csv(sim_file, show_col_types = FALSE)
  cat("Loaded existing simulation results.\n")
}
```

# SIMULATED DATA
```{r}
sim_data <- read_csv(glue::glue("{main_path}simdata/simulated_responses.csv"))
```

# STAN MODEL COMPILATION
```{r}
stan_file_gcm_single <- "stan/ch12_gcm_single.stan"
mod_gcm_single <- cmdstan_model(stan_file_gcm_single)

```


# FIT FOR SINGLE AGENT
```{r}
agent_to_fit <- sim_data |>
  filter(
    round(w4_true, 1) == 0.6,
    abs(log_c_true - log(2.0)) == min(abs(log_c_true - log(2.0)))
  ) |>
  filter(agent_id == min(agent_id)) |>
  slice(1:total_trials)

stopifnot(nrow(agent_to_fit) == total_trials)

gcm_data_single <- list(
  N_total           = nrow(agent_to_fit),
  N_features        = 5L,
  y                 = as.integer(agent_to_fit$sim_response),
  obs               = as.matrix(agent_to_fit[, c("feature1", "feature2", "feature3", "feature4", "feature5")]),
  danger_feedback      = as.integer(agent_to_fit$danger_feedback),
  prior_w_alpha     = rep(1, 5),
  prior_log_c_mu    = log(2), 
  prior_log_c_sigma = 1, 
  prior_bias_alpha  = 1,
  prior_bias_beta   = 1,
  run_diagnostics   = 1L
)

# SINGLE AGENT FIT PATH
fit_filepath_single <- glue::glue("{main_path}src/stan/ch12_gcm_single_fit.rds")

if (regenerate_simulations || !file.exists(fit_filepath_single)) {
  pf_single <- tryCatch(
    mod_gcm_single$pathfinder(
      data      = gcm_data_single,
      seed      = 123,
      num_paths = 4,
      refresh   = 0
    ),
    error = function(e) {
      message("Pathfinder failed — falling back to default initialisation.")
      NULL
    }
  )

  fit_gcm_single <- mod_gcm_single$sample(
    data            = gcm_data_single,
    init            = pf_single,
    seed            = 123,
    chains          = 4,
    parallel_chains = min(4, availableCores()),
    iter_warmup     = 500,
    iter_sampling   = 1500,
    refresh         = 300,
    adapt_delta     = 0.9
  )
  fit_gcm_single$save_object(fit_filepath_single)
  cat("Single-agent model fitted and saved.\n")
} else {
  fit_gcm_single <- readRDS(fit_filepath_single)
  cat("Loaded existing single-agent fit.\n")
}
```

READ SINGLE AGENT FIT IN 
```{r}
readRDS(glue::glue("{main_path}src/stan/ch12_gcm_single_fit.rds"))

```


MODEL FIT - MCMC Diagnostics

```{r}
# Trace and Rank plots
bayesplot::mcmc_trace(
    fit_gcm_single$draws(c("log_c", "w[4]", "bias")),
    facet_args = list(ncol = 1)
  )

bayesplot::mcmc_rank_overlay(
    fit_gcm_single$draws(c("log_c", "w[4]", "bias"))
  )
```

# Check for Divergent Transitions - Chain level diagnostics
```{r}
fit_gcm_single$diagnostic_summary()
```

Here we can see that no chains diverged, and the minimum E-BFMI (Energy Bayesian Fraction of Missing Information) > 0.3, which is good. The sampler never hit the maximum allowed tree depth during sampling which means that it stopped naturally and it is working efficiently.


# Check R-hat, EBFMI and ESS metrics
```{r}
fit_summary <- fit_gcm_single$summary(variables = c("w[1]", "w[2]", "w[3]", "w[4]", "w[5]", "log_c", "c", "bias"))
fit_summary <- as.data.frame(fit_summary)

print(fit_summary %>% select(variable, mean, rhat, ess_bulk, ess_tail))
```

The R-hat values are all < 1.1, which is good. And the ess bulk and tail are both > 1000, meaning there are effective samples in the center and tails of the distribution. Overall, the MCMC diagnostics battery shows good convergence and  sampling.

## Model Checks - single subject

PRIOR PREDICTIVE CHECK
```{r}
n_prior_samples <- 500

# Define prior distributions for the parameters
prior_draws <- tibble(
  sample_id = seq_len(n_prior_samples),
  log_c = rnorm(n_prior_samples, log(2), 1),
  w4 = runif(n_prior_samples),
  bias = rbeta(n_prior_samples, 1, 1)
) %>% 
  mutate( c_true = exp(log_c),
          w_remaining = (1-w4) / 4,
          w1 = w_remaining,
          w2 = w_remaining,
          w3 = w_remaining,
          w5 = w_remaining)

ppc_schedule <- make_subject_schedule(sim_grid, n_blocks, seed = 123)


# prior predictive curves
prior_pred_curves <- pmap_dfr(
  list(
    sample_id = prior_draws$sample_id, # sample ID
    log_c = prior_draws$log_c, # log sensitivity
    w4 = prior_draws$w4, # attention weight on feature 4
    bias = prior_draws$bias # response bias
  ),
  
  function(sample_id, log_c, w4, bias){
    c <- exp(log_c) # convert log sensitivity back to natural scale
    sim <- gcm_simulate(
      w = c((1-w4)/4, (1-w4)/4, (1-w4)/4, w4, (1-w4)/4), # attention weights
      c = c, # sensitivity
      bias = bias, # response bias
      features = ppc_schedule |> dplyr::select(feature1, feature2, feature3, feature4, feature5), # stimulus features
      danger = ppc_schedule$danger_feedback # true category labels
    )
    tibble(
      sample_id = sample_id,
      trial = seq_len(nrow(ppc_schedule)),
      correct = as.integer(sim$sim_response == ppc_schedule$danger_feedback)
    ) %>% 
      mutate(cum_acc = cumsum(correct) / row_number())
  }
)

# summarise prior predictive curves to get quantiles of cumulative accuracy across trials
prior_pred_summary <- prior_pred_curves %>% 
  group_by(trial) %>% 
  summarise(
    q05 = quantile(cum_acc, 0.05),
    q25 = quantile(cum_acc, 0.25),
    q50 = median(cum_acc),
    q75 = quantile(cum_acc, 0.75),
    q95 = quantile(cum_acc, 0.95),
    .groups = "drop"
  )

# plot preior predictive check
ggplot(prior_pred_summary, aes(x = trial)) +
  geom_ribbon(aes(ymin = q05, ymax = q95), fill = "#56B4E9", alpha = 0.20) +
  geom_ribbon(aes(ymin = q25, ymax = q75), fill = "#56B4E9", alpha = 0.40) +
  geom_line(aes(y = q50), color = "#0072B2", linewidth = 1) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey40") +
  scale_y_continuous(limits = c(0, 1)) +
  labs(
    title    = "Prior Predictive Check: GCM Learning Curves",
    subtitle = "Ribbons: 50% and 90% prior predictive intervals; line: median\nPriors: log(c) ~ Normal(log 2, 1),  w ~ Dirichlet(1,1),  bias ~ Beta(1,1)",
    x = "Trial", y = "Cumulative Accuracy"
  )
```

POSTERIOR PREDICTIVE CHECK
```{r}
if (!is.null(fit_gcm_single)) {
  prob_draws <- fit_gcm_single$draws("prob_danger1", format = "matrix")
  n_draws    <- nrow(prob_draws)
  n_trials   <- ncol(prob_draws)

  set.seed(2025)
  yrep_post <- matrix(rbinom(n_draws * n_trials, 1, as.vector(prob_draws)),
                      nrow = n_draws)

  obs_acc <- cummean(agent_to_fit$sim_response == agent_to_fit$danger_feedback)
  rep_acc <- t(apply(yrep_post, 1, function(yr) {
    cummean(yr == agent_to_fit$danger_feedback)
  }))

  ppc_summary <- tibble(
    trial = seq_len(n_trials),
    obs   = obs_acc,
    q05   = apply(rep_acc, 2, quantile, 0.05),
    q25   = apply(rep_acc, 2, quantile, 0.25),
    q50   = apply(rep_acc, 2, quantile, 0.50),
    q75   = apply(rep_acc, 2, quantile, 0.75),
    q95   = apply(rep_acc, 2, quantile, 0.95)
  )

  ggplot(ppc_summary, aes(x = trial)) +
    geom_ribbon(aes(ymin = q05, ymax = q95), fill = "#56B4E9", alpha = 0.20) +
    geom_ribbon(aes(ymin = q25, ymax = q75), fill = "#56B4E9", alpha = 0.40) +
    geom_line(aes(y = q50), color = "#0072B2", linewidth = 0.8) +
    geom_line(aes(y = obs), color = "#D55E00", linewidth = 1) +
    scale_y_continuous(limits = c(0, 1)) +
    labs(
      title    = "Posterior Predictive Check: Cumulative Accuracy",
      subtitle = "Blue ribbons: 50% and 90% posterior predictive intervals\nOrange line: observed cumulative accuracy",
      x = "Trial", y = "Cumulative Accuracy"
    )
}
```
 The cumulative accuracy is chosen for the PPC as it provides a better overview of the learning dynamics, compared to trial-level accuracy. The GCM's path dependence means that trial-level predictions are more variable and noisy, making it less insightful about model fit.


PRIOR POSTERIOR UPDATE
```{r}
if (!is.null(fit_gcm_single)) {
  draws_df <- as_draws_df(
    fit_gcm_single$draws(variables = c("w[4]", "log_c", "c", "bias"))
  )

  true_w4    <- agent_to_fit$w4_true[4]
  true_log_c <- agent_to_fit$log_c_true[1]
  true_c     <- agent_to_fit$c_true[1]
  true_bias  <- agent_to_fit$bias_true[1]

  set.seed(2026)
  n_prior <- 4000
  prior_df <- tibble(
    `w[4]` = runif(n_prior),
    log_c  = rnorm(n_prior, log(2), 1),
    c      = exp(log_c),
    bias   = rbeta(n_prior, 1, 1)
  )

  pp_panel <- function(prior_x, post_x, true_val, label) {
    tibble(
      x      = c(prior_x, post_x),
      source = c(rep("Prior", length(prior_x)),
                 rep("Posterior", length(post_x)))
    ) |>
      ggplot(aes(x = x, fill = source, color = source)) +
      geom_density(alpha = 0.4, linewidth = 0.5) +
      geom_vline(xintercept = true_val, color = "red",
                 linetype = "dashed", linewidth = 0.7) +
      scale_fill_manual(values = c(Prior = "grey60", Posterior = "#0072B2")) +
      scale_color_manual(values = c(Prior = "grey40", Posterior = "#0072B2")) +
      labs(title = label, x = NULL, y = "Density") +
      theme(legend.position = "bottom")
  }

  p1 <- pp_panel(prior_df$`w[4]`, draws_df$`w[4]`, true_w4, "w[4]")
  p2 <- pp_panel(prior_df$log_c,  draws_df$log_c,  true_log_c, "log(c) — computational scale")
  p3 <- pp_panel(prior_df$c,      draws_df$c,      true_c,     "c — natural scale")
  p4 <- pp_panel(prior_df$bias,   draws_df$bias,   true_bias,  "bias")

  (p1 | p2) / (p3 | p4) +
    plot_annotation(
      title    = "Prior–Posterior Update for the Single-Subject GCM",
      subtitle = "Red dashed line = true generating value"
    )
}
```

JOINT POSTERIOR GEOMETRY - Check for no correlation between parameters
```{r}
if (!is.null(fit_gcm_single)) {
  bayesplot::mcmc_pairs(
    fit_gcm_single$draws(c("log_c", "w[4]", "bias")),
    diag_fun     = "dens",
    off_diag_fun = "hex"
  )
}
```

LOO with PARETO-K DIAGNOSTICS - Check for issues with influential data points
```{r}
if (!is.null(fit_gcm_single)) {
  log_lik_mat    <- fit_gcm_single$draws("log_lik", format = "matrix")
  loo_gcm_single <- loo::loo(log_lik_mat, cores = 4)
  print(loo_gcm_single)

  plot(loo_gcm_single, label_points = TRUE,
       main = "PSIS-LOO Pareto-k diagnostics: Single-subject GCM")
}
```
Pareto k estimates are good if k < 0.7.

# PLOTS

PLOTS PREP
```{r}
# Load sim_data
sim_data <- read_csv(glue::glue("{main_path}/simdata/simulated_responses.csv"))
 

# ADD LABELS
sim_labeled <- sim_data %>%
  mutate(
    c_label  = fct_reorder(paste0("c = ", round(c_true, 1)), c_true),
    w4_label = case_when(
      w4_true == 0.2 ~ "w4 = 0.2 (uniform)",
      w4_true == 1.0 ~ "w4 = 1.0 (extreme)",
      w4_true == 0.6 ~ "w4 = 0.6 (middle)",
      w4_true == 0.8 ~ "w4 = 0.8 (strong)",
      TRUE           ~ paste0("w4 = ", w4_true)
    ),
    w4_label = fct_reorder(w4_label, w4_true)
  )
 
# Block boundary positions
block_boundaries <- sim_labeled %>%
  group_by(block) %>%
  summarise(end_trial = max(trial_within_subject), .groups = "drop") %>%
  filter(block < max(block)) %>%
  pull(end_trial) %>%
  {. + 0.5}
```
ACTUAL PLOTS TO COMPARE SIMUALTED DATA
```{r}
# CUMULATIVE ACCURACY — effect of c, faceted by w4
p_c_effect <- ggplot(
  sim_labeled,
  aes(x      = trial_within_subject,
      y      = performance,
      colour = c_label,
      group  = interaction(agent_id, c_label))
) +
  stat_summary(fun = mean, geom = "line",
               aes(group = c_label), linewidth = 1.1) +
  geom_vline(xintercept = block_boundaries,
             linetype = "dashed", alpha = 0.4, colour = "grey30") +
  facet_wrap(~w4_label) +
  scale_colour_viridis_d(option = "plasma", name = "Sensitivity (c)") +
  labs(
    title    = "Effect of sensitivity (c) on GCM performance",
    subtitle = "Mean cumulative accuracy across agents; dashed = block boundary",
    x = "Trial", y = "Cumulative accuracy"
  ) +
  ylim(0, 1) +
  theme_minimal(base_size = 12) +
  theme(legend.position  = "bottom",
        strip.text       = element_text(face = "bold"),
        panel.grid.minor = element_blank())
 

# CUMULATIVE ACCURACY — effect of w4, faceted by c
p_w_effect <- ggplot(
  sim_labeled,
  aes(x      = trial_within_subject,
      y      = performance,
      colour = w4_label,
      group  = interaction(agent_id, w4_label))
) +
  stat_summary(fun = mean, geom = "line",
               aes(group = w4_label), linewidth = 1.1) +
  geom_vline(xintercept = block_boundaries,
             linetype = "dashed", alpha = 0.4, colour = "grey30") +
  facet_wrap(~c_label) +
  scale_colour_brewer(palette = "Set1", name = "Attention to feature 4 (w4)") +
  labs(
    title    = "Effect of attention weight on feature 4 (w4) on GCM performance",
    subtitle = "Mean cumulative accuracy across agents; dashed = block boundary",
    x = "Trial", y = "Cumulative accuracy"
  ) +
  ylim(0, 1) +
  theme_minimal(base_size = 12) +
  theme(legend.position  = "bottom",
        strip.text       = element_text(face = "bold"),
        panel.grid.minor = element_blank())
 

# ROLLING ACCURACY (window = 8)
roll_k <- function(x, k = 8) as.numeric(stats::filter(x, rep(1/k, k), sides = 1))
 
rolling <- sim_labeled %>%
  arrange(agent_id, trial_within_subject) %>%
  group_by(agent_id, c_label, w4_label) %>%
  mutate(rolling_acc = roll_k(correct, k = 8)) %>%
  ungroup()
 
p_rolling <- ggplot(
  rolling,
  aes(x      = trial_within_subject,
      y      = rolling_acc,
      colour = c_label,
      group  = interaction(agent_id, c_label))
) +
  stat_summary(fun = mean, geom = "line",
               aes(group = c_label), linewidth = 1.1, na.rm = TRUE) +
  geom_vline(xintercept = block_boundaries,
             linetype = "dashed", alpha = 0.4, colour = "grey30") +
  facet_wrap(~w4_label) +
  scale_colour_viridis_d(option = "plasma", name = "Sensitivity (c)") +
  labs(
    title    = "Rolling accuracy (window = 8 trials)",
    subtitle = "Captures within-block learning; dashed = block boundary",
    x = "Trial", y = "Rolling accuracy"
  ) +
  ylim(0, 1) +
  theme_minimal(base_size = 12) +
  theme(legend.position  = "bottom",
        strip.text       = element_text(face = "bold"),
        panel.grid.minor = element_blank())
 

# P(dangerous) OVER TRIALS — model confidence by true label
p_prob <- ggplot(
  sim_labeled,
  aes(x      = trial_within_subject,
      y      = prob_dan1,
      colour = factor(danger_feedback))
) +
  geom_smooth(se = FALSE, linewidth = 0.9, method = "loess", span = 0.3) +
  geom_vline(xintercept = block_boundaries,
             linetype = "dashed", alpha = 0.4, colour = "grey30") +
  facet_wrap(~c_label) +
  scale_colour_manual(
    values = c("0" = "#0072B2", "1" = "#D55E00"),
    labels = c("Safe", "Dangerous"),
    name   = "True label"
  ) +
  labs(
    title    = "Model confidence P(dangerous) by true category",
    subtitle = "Loess-smoothed; separation between lines = discrimination",
    x = "Trial", y = "P(dangerous)"
  ) +
  ylim(0, 1) +
  theme_minimal(base_size = 12) +
  theme(legend.position  = "bottom",
        strip.text       = element_text(face = "bold"),
        panel.grid.minor = element_blank())
 
# Display plots
print(p_c_effect)
print(p_w_effect)
print(p_rolling)
print(p_prob)
```
#Comments on the results:
plot 1
Effect of c, faceted by w4
The ceiling performance improves as w4 increases from uniform (0.2) to extreme (1.0), but only for high c values. Yellow (c=15) reaches ~90% with uniform attention but nearly 100% when w4=0.6 or higher — meaning sensitivity alone gets you far, but directing attention to the right feature (arms up, the session-2 danger rule) pushes it further. Crucially, low c (dark blue) barely changes across facets — when you're not sensitive enough to exploit fine-grained similarity, it doesn't matter where you direct your attention.

plot 2
Effect of w4, faceted by c
This is the clearest story. At low c (top row), all four w4 settings produce nearly identical curves — the lines cluster together. At high c (bottom row), the split becomes dramatic: red (w4=0.2, uniform) plateaus around 60–65% while purple (w4=1.0, extreme) rockets to near-perfect accuracy very quickly. This is the key model prediction: w4 and c interact — attention only matters when sensitivity is high enough to act on it. Neither parameter alone is sufficient for good performance.

plot 3
Rolling accuracy
Compared to the previous version, the within-block structure is now much cleaner. A few things stand out: the block boundaries (dashed lines) don't show visible dips like before — this makes sense because you're only fitting session 2, so the category rule is stable across all 3 blocks. The main remaining variability is just noise from the small window size. Yellow (c=15) is consistently at or near 1.0 from early on regardless of w4, while low-c agents stay noisy throughout.

plot 4
At c=0.5, the orange (truly dangerous) and blue (safe) lines barely separate — the model essentially guesses. As c increases, the gap widens dramatically: by c=15, orange is near 1.0 and blue near 0.0. Notice also that the separation is established very early (within the first ~10 trials) and then stays flat — the model isn't gradually learning across blocks, it locks in quickly. The block boundaries (dashed) cause no disruption at all, confirming that within-session rule stability means memory from earlier blocks helps rather than hurts.

Thus, the two parameters play distinct roles — c governs how sharply the model discriminates at all, and w4 governs how efficiently it uses the most diagnostic feature. Their interaction is the core finding: high w4 without high c is wasted, and high c with low w4 hits a ceiling around 65% that it can't break through.


# EXPERIMENTAL DATA

LOADING EXPERIMENTAL DATA
```{r}
#Loading the exp_data
exp_data <- read_csv(glue::glue("{main_path}/data/AlienData.txt"))
```

PLOT PREP FOR COMPARING SIMULATED AND EMPIRICAL DATA
```{r}
# Empirical mean per participant — session 2, individual condition only
emp_by_subject <- exp_data %>%
  filter(
    session == 2,
    condition == 2 
  ) %>%
  group_by(subject) %>%
  summarise(
    m  = mean(correct, na.rm = TRUE),
    se = sd(correct, na.rm = TRUE) / sqrt(sum(!is.na(correct))),
    .groups = "drop"
  ) %>%
  mutate(
    label = paste0("S", subject),
    type  = "Empirical subjects"
  )

# Simulated mean per agent type
sim_by_type <- sim_data %>%
  mutate(
    label = paste0("c=", round(c_true, 1), ", w4=", round(w4_true, 1))
  ) %>%
  group_by(label, c_true, w4_true) %>%
  summarise(
    m  = mean(correct, na.rm = TRUE),
    se = sd(correct, na.rm = TRUE) / sqrt(sum(!is.na(correct))),
    .groups = "drop"
  ) %>%
  mutate(
    type = "Simulated agent types"
  )

# COMBINE & PLOT SIDE BY SIDE
combined <- bind_rows(
  emp_by_subject %>% select(label, m, se, type),
  sim_by_type %>% select(label, m, se, type)
)

p_comparison <- ggplot(combined, aes(x = reorder(label, m), y = m, fill = type)) +
  geom_col(width = 0.7, alpha = 0.85) +
  geom_errorbar(aes(ymin = m - se, ymax = m + se), width = 0.25) +
  geom_hline(yintercept = 0.5, linetype = "dashed", alpha = 0.4) +
  facet_wrap(~ type, scales = "free_x") +
  scale_y_continuous(limits = c(0, 1), breaks = seq(0, 1, 0.25)) +
  labs(
    title = "Mean accuracy: Empirical subjects vs. simulated agent types",
    subtitle = "Empirical (individual condition); dashed line = chance (0.5)",
    x = NULL,
    y = "Proportion correct"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "none",
    axis.text.x = element_text(
      size = 8,
      angle = 90,   # to have the bar labels vertically
      vjust = 0.5,
      hjust = 1
    ),
    panel.grid.minor = element_blank()
  )

print(p_comparison)
```

PLOTTING EMPIRICAL DATA
```{r}
# plotting empirical data, individual condition only, excluding the first 8 training trials
exp_data <- read_csv(glue::glue("{main_path}/data/AlienData.txt")) 

d <- exp_data %>%
  filter(condition == 2) %>%   # Individual condition only
  mutate(
    danger_resp    = as.integer(response %in% c(3, 4)),
    nutri_resp     = as.integer(response %in% c(2, 4)),
    correct_danger = as.integer(danger_resp == dangerous),
    correct_nutri  = as.integer(nutri_resp  == nutricious),
    correct_both   = as.integer(correct_danger == 1 & correct_nutri == 1),
    session_f      = factor(session, labels = c("Session 1", "Session 2", "Session 3")),
    condition_f    = "Individual"
  ) %>%
  group_by(subject, session) %>%
  arrange(cycle, trial, .by_group = TRUE) %>%
  mutate(original_trial_in_session = row_number()) %>%
  filter(original_trial_in_session > 8) %>%   # to remove first 8 training trials
  mutate(trial_in_session = row_number()) %>% # to renumber remaining 96 trials
  ungroup() %>%
  group_by(subject, session) %>%
  arrange(trial_in_session, .by_group = TRUE) %>%
  mutate(
    trial_global = row_number(),
    performance  = cumsum(correct_danger) / row_number()
  ) %>%
  ungroup()

block_boundaries_sess <- c(32.5, 64.5)

theme_set(theme_minimal(base_size = 12))
pal_cond <- c("Individual" = "#D55E00")

# CUMULATIVE ACCURACY OVER TRIALS
p_cumulative <- d %>%
  group_by(subject, session_f) %>%
  arrange(trial_in_session, .by_group = TRUE) %>%
  mutate(cumulative = cumsum(correct_danger) / row_number()) %>%
  ungroup() %>%
  group_by(condition_f, session_f, trial_in_session) %>%
  summarise(
    m  = mean(cumulative, na.rm = TRUE),
    se = sd(cumulative, na.rm = TRUE) / sqrt(sum(!is.na(cumulative))),
    .groups = "drop"
  ) %>%
  ggplot(aes(trial_in_session, m,
             colour = condition_f, fill = condition_f)) +
  geom_ribbon(aes(ymin = m - se, ymax = m + se),
              alpha = 0.15, colour = NA) +
  geom_line(linewidth = 0.9) +
  geom_vline(xintercept = block_boundaries_sess,
             linetype = "dashed", alpha = 0.4, colour = "grey30") +
  facet_wrap(~session_f) +
  scale_colour_manual(values = pal_cond, name = "Condition") +
  scale_fill_manual(values = pal_cond, name = "Condition") +
  labs(
    title = "Empirical: Cumulative accuracy over trials (danger dimension)",
    subtitle = "Individual condition only; first 8 training trials removed; Mean ± SE",
    x = "Trial within session",
    y = "Cumulative accuracy"
  ) +
  ylim(0, 1) +
  theme(
    legend.position = "bottom",
    strip.text = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )

# ROLLING ACCURACY
roll_k <- function(x, k = 8) {
  as.numeric(stats::filter(x, rep(1/k, k), sides = 1))
}

p_rolling <- d %>%
  arrange(subject, session, trial_in_session) %>%
  group_by(subject, session_f, condition_f) %>%
  mutate(rolling = roll_k(correct_danger, k = 8)) %>%
  ungroup() %>%
  group_by(condition_f, session_f, trial_in_session) %>%
  summarise(
    m  = mean(rolling, na.rm = TRUE),
    se = sd(rolling, na.rm = TRUE) / sqrt(sum(!is.na(rolling))),
    .groups = "drop"
  ) %>%
  ggplot(aes(trial_in_session, m,
             colour = condition_f, fill = condition_f)) +
  geom_ribbon(aes(ymin = m - se, ymax = m + se),
              alpha = 0.15, colour = NA) +
  geom_line(linewidth = 0.9) +
  geom_vline(xintercept = block_boundaries_sess,
             linetype = "dashed", alpha = 0.4, colour = "grey30") +
  facet_wrap(~session_f) +
  scale_colour_manual(values = pal_cond, name = "Condition") +
  scale_fill_manual(values = pal_cond, name = "Condition") +
  labs(
    title = "Empirical: Rolling accuracy (window = 8, danger dimension)",
    subtitle = "Individual condition only; first 8 training trials removed; Mean ± SE",
    x = "Trial within session",
    y = "Rolling accuracy"
  ) +
  ylim(0, 1) +
  theme(
    legend.position = "bottom",
    strip.text = element_text(face = "bold"),
    panel.grid.minor = element_blank()
  )

print(p_cumulative)
print(p_rolling)
```


