---
title: "Script"
output: html_document
---

# Setup
```{r}
pacman::p_load(
  tidyverse,    # Data manipulation and plotting
  here,         # Robust file paths
  cmdstanr,     # Stan interface
  rstan,
  posterior,    # Posterior analysis
  bayesplot,    # Plotting MCMC
  furrr,        # Parallel processing
  future,        # Future backend
  patchwork
)
```

```{r}
# Parallel Setup: Leave 2 cores free for system stability
n_cores <- parallel::detectCores() - 2
plan(multisession, workers = n_cores)
```

```{r}
# Optional: Set a default ggplot theme
theme_set(theme_classic())
```

# Simulating Data
```{r}
# Flag to control whether to regenerate simulation/fitting results for the knitted files
# Set to TRUE to rerun everything (takes time!), FALSE to load saved results.
regenerate_simulations <- TRUE 

```

# Define agents.
The two agents are: WBA and SBA

The variables:
ID : Participant ID
1_rating: The first rating the participant gives (0-8)
g_rating: The group rating feedback given after the first rating (0-8)
2_rating: The second rating the participant gives (0-8) <- this is what we want to predict
Feedback: How far is the g_rating from the 1_rating (-3 to 3)
Change: How much did the 2_rating change from the 1_rating (-7 to 7) but probably in the middle
Condition: Pre or Peri depending on whether the data was acquired before or during the pandemic

We can delete any rows with NAs!!

Social Conformity Experiment

 Goal: simulate agents that combine their own rating (FirstRating) with
       a group rating (GroupRating) to produce a second rating (SecondRating).

 Scale: ratings are integers 1–8 (1 = not trustworthy, 8 = very trustworthy)

 Key adaptation from the chapter's marble task:
   Marble task  → binary choice,   counts k out of n
   This task    → ordinal rating,  integers 1–8
   We convert ratings to pseudo-counts:
     k = rating - 1   (so rating 1 → 0 "blue", rating 8 → 7 "blue")
     n = 7            (total pseudo-marbles per source)
   This puts both sources on the same 0–7 count scale.
   Belief (theta) lives on [0, 1] and maps back:
     SecondRating = round(theta * 7) + 1
     
     
```{r}
library(tidyverse)
set.seed(42)
```

```{r}
# Rating scale boundaries
RATING_MIN <- 1
RATING_MAX <- 8
N_PSEUDO   <- 7   # pseudo sample size per source (= RATING_MAX - RATING_MIN)
 
# Jeffreys prior
alpha_prior <- 0.5
beta_prior <- 0.5
 

# Convert a Beta belief (0–1) back to a 1–8 rating (rounded, clamped)

```

```{r}
# SECTION 2: Core belief-update function (shared by all agents)
#
# Implements Bayesian updating on a Beta distribution:
#   posterior = Beta(alpha_prior + w_self*k_self + w_group*k_group,
#                   beta_prior  + w_self*(n-k_self) + w_group*(n-k_group))
#
# Arguments:
#   rating_1  – participant's own rating (integer 1–8)
#   rating_g  – group's rating (integer 1–8)
#   w_self        – weight on own evidence  (WBA/PBA: free)
#   w_group       – weight on group evidence (WBA/PBA: free)

compute_posterior <- function(rating_1, rating_g,
                              w_self = 1, w_group = 1) {
  k_self  <- rating_1 - 1 #converting ratings (1–8) to pseudo-counts (0–7)
  k_group <- rating_g - 1 
 
  alpha_post <- alpha_prior + w_self * k_self + w_group * k_group
  beta_post  <- beta_prior  + w_self * (N_PSEUDO - k_self) +
                               w_group * (N_PSEUDO - k_group)
 
  list(
    alpha = alpha_post,
    beta  = beta_post,
    expected_rate  = alpha_post / (alpha_post + beta_post)
  )
}
```

Three agent functions

 Each agent takes (first_rating, group_rating) and returns a SecondRating.
 All use probability matching:
   1. Sample theta ~ Beta(alpha_post, beta_post)   [belief about trustworthiness]
   2. SecondRating = round(theta * 7) + 1          [express as a 1–8 rating]

```{r}
# --- Agent B: Weighted Bayesian Agent (WBA) -----------------------------------
# Allows different weights for own vs. group evidence.
# w_self and w_group are both free parameters (>0).
 
wba_agent <- function(rating_1, rating_g, rho, kappa) {
  
  w_self <- rho * kappa 
  w_group <- (1 - rho) * kappa
  
  # convert sample to count and compute posterior
  post <- compute_posterior(rating_1, rating_g, w_self, w_group)
  
  # sample from the posterior belief 
  theta <- rbeta(1, post$alpha, post$beta)
  
  #convert back to a rating
  rating_2 <- round(theta * N_PSEUDO) + 1
  
  return(list(rating_1 = rating_1, rating_2 = rating_2, w_self = w_self, w_group = w_group))
}
```

```{r}
# --- Agent C: Proportional Bayesian Agent (PBA) ------------------------------
# Reparameterises WBA to a single identifiable weight w ∈ (0, 1):
#   w     = proportion of total weight given to own evidence
#   1 - w = proportion given to group evidence
#
# This fixes the identifiability problem:
#   w = 0.5  → equal weights (identical to SBA in belief direction,
#               but overall precision differs because total weight = 1, not 2)
#   w > 0.5  → trusts self more   (under-conformist)
#   w < 0.5  → trusts group more  (over-conformist / conformist)
 
pba_agent <- function(rating_1, rating_g, w) {
  
  # convert sample to count and compute posterior with proportional weights
  post  <- compute_posterior(rating_1, rating_g, w_self = w, w_group = 1 - w)
  
  # sample from the posterior belief
  theta <- rbeta(1, post$alpha, post$beta)
  
  # convert back to a rating
  rating_2 <- round(theta * N_PSEUDO) + 1
  
  return(list(rating_1 = rating_1, rating_2 = rating_2, w = w))
}
```

# Simulate the experiment 
```{r}
simulate_experiment <- function(agent_fn,
                                agent_args = list(),
                                n_faces    = 5) {
  
  offsets <- c(-3, -2, 2, 3)
  results <- vector("list", n_faces)
  idx <- 1
  
  for (face_id in 1:n_faces) {
      
      # Sample ratings
      rating_1 <- sample(1:8, 1)
      valid_offsets <- offsets[rating_1 + offsets >= 1 & rating_1 + offsets <= 8]
      feedback <- sample(valid_offsets, 1)
      rating_g <- rating_1 + feedback
      
      # Call agent
      result <- do.call(agent_fn, c(list(rating_1 = rating_1, rating_g = rating_g), agent_args))
      
      results[[idx]] <- tibble(
        FaceID       = face_id,
        FirstRating  = rating_1,
        GroupRating  = rating_g,
        SecondRating = result$rating_2,
        Feedback     = feedback,
        Change       = result$rating_2 - rating_1,
        total_g      = 8,
        total_1      = 8,
        N            = n_faces
      ) |> bind_cols(as_tibble(agent_args))
      
      idx <- idx + 1
  } 
  
  bind_rows(results)
}
```

3. Design 3–5 distinguishing scenarios
Pick parameter combinations that make SBA and WBA/PBA behave differently. E.g.:

Equal weights (SBA = WBA)
Strong self-trust (WBA with high w_self)
Strong conformity (WBA with high w_group)
Ignoring social info completely
Ignoring own prior completely

Scenarios for WBA and PBA:
SCENARIO 01: (works only for wba - strong conformity)
kappa = 11
rho = 0.09
wd = 1
ws = 10

SCENARIO 02: (works for both models - balanced)
kappa = 1
rho = 0.5
wd = 0.5
ws= 0.5
 
SCENARIO 03: (works for both models - self focused)
kappa = 1
rho = 0.9
wd = 0.9
ws = 0.1



## WBA scenarios

### WBA: Strong conformity
```{r}
sim_df_wba_social <- simulate_experiment(wba_agent, agent_args = list(rho = 0.09, kappa = 11), n_faces = 20)
```

### WBA: Balanced
```{r}
sim_df_wba_balanced <- simulate_experiment(wba_agent, agent_args = list(rho = 0.5, kappa = 1), n_faces = 20)

```

### WBA: Self focused
```{r}
sim_df_wba_self <- simulate_experiment(wba_agent, agent_args = list(rho = 0.9, kappa = 1), n_faces = 20)

```

## PBA scenarios 

### PBA: Strong conformity...
```{r}
sim_df_pba_social <- simulate_experiment(pba_agent, agent_args = list(w = 0.1), n_faces = 20)

```

### PBA: Balanced
```{r}
sim_df_pba_balanced <- simulate_experiment(pba_agent, agent_args = list(w = 0.5), n_faces = 20)

```

### PBA: Self Focused
```{r}
sim_df_pba_self <- simulate_experiment(pba_agent, agent_args = list(w = 0.9), n_faces = 20)

```


```{r}
# combine and label your scenarios
sim_pba_all <- bind_rows(
  simulate_experiment(pba_agent, list(w = 0.1), n_faces = 200) |> mutate(scenario = "Conformist (w=0.1)"),
  simulate_experiment(pba_agent, list(w = 0.5), n_faces = 200) |> mutate(scenario = "Balanced (w=0.5)"),
  simulate_experiment(pba_agent, list(w = 0.9), n_faces = 200) |> mutate(scenario = "Self-focused (w=0.9)")
)

# Then plot
ggplot(sim_pba_all, aes(x = Feedback, y = Change, color = scenario)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm") +
  facet_wrap(~scenario) +
  labs(title = "PBA: Change as a function of Group Feedback",
       x = "Feedback (GroupRating - FirstRating)",
       y = "Change (SecondRating - FirstRating)")
```
### Model Fitting 
```{r}
mod_simple       <- cmdstan_model("/Users/magnusseverin/Documents/GitHub/Advanced_Cognitive_Modeling/Assignment03/src/ProbAgent.stan")
#mod_weighted_raw <- cmdstan_model("/Users/magnusseverin/Documents/GitHub/Advanced_Cognitive_Modeling/Assignment03/src/WeightedAgent.stan")

sampler_args <- list(
  chains = 4, 
  parallel_chains = 4, 
  iter_warmup = 1000, 
  iter_sampling = 1000, 
  refresh = 0
)
```



```{r}
stan_data_pba <- list(
  balanced = as.list(sim_df_pba_balanced),
  self     = as.list(sim_df_pba_self),
  social   = as.list(sim_df_pba_social)
)
```

```{r}
stan_data_pba$balanced$N <- as.integer(stan_data_pba$balanced$N[[1]])
stan_data_pba$self$N     <- as.integer(stan_data_pba$self$N[[1]])
stan_data_pba$social$N   <- as.integer(stan_data_pba$social$N[[1]])
```

```{r}
fit_pba_paths <- here::here(
  "/Users/magnusseverin/Documents/GitHub/Advanced_Cognitive_Modeling/Assignment03/src/simmodels",
  paste0("fit_pba_", names(stan_data_pba), ".rds")
) |> set_names(names(stan_data_pba))

if (regenerate_simulations || !all(file.exists(fit_pba_paths))) {
  fit_pba <- imap(stan_data_pba, function(data, scenario) {
    fit <- mod_simple$sample(
      data          = data,
      chains        = 2,
      iter_warmup   = 0,
      iter_sampling = 1000,
      fixed_param   = TRUE,
      refresh       = 0
    )
    fit$save_object(fit_pba_paths[[scenario]])
    fit
  })
} else {
  fit_pba <- map(fit_pba_paths, readRDS)
}
```