---
title: "Script"
output: html_document
---

# Setup
```{r}
pacman::p_load(
  tidyverse,    # Data manipulation and plotting
  here,         # Robust file paths
  cmdstanr,     # Stan interface
  rstan,
  posterior,    # Posterior analysis
  bayesplot,    # Plotting MCMC
  furrr,        # Parallel processing
  future,        # Future backend
  patchwork
)
```

```{r}
# Parallel Setup: Leave 2 cores free for system stability
n_cores <- parallel::detectCores() - 2
plan(multisession, workers = n_cores)
```

```{r}
# Optional: Set a default ggplot theme
theme_set(theme_classic())
```

# Simulating Data
```{r}
# Flag to control whether to regenerate simulation/fitting results for the knitted files
# Set to TRUE to rerun everything (takes time!), FALSE to load saved results.
regenerate_simulations <- TRUE # Or TRUE

```

# Define agents.
The two agents are: WBA and SBA

The variables:
ID : Participant ID
1_rating: The first rating the participant gives (0-8)
g_rating: The group rating feedback given after the first rating (0-8)
2_rating: The second rating the participant gives (0-8) <- this is what we want to predict
Feedback: How far is the g_rating from the 1_rating (-3 to 3)
Change: How much did the 2_rating change from the 1_rating (-7 to 7) but probably in the middle
Condition: Pre or Peri depending on whether the data was acquired before or during the pandemic

We can delete any rows with NAs!!


```{r}
# AGENT SBA

```

```{r}
# AGENT WBA

```


# Simulate the task
```{r}

```


```{r}

```

