---
title: "Script"
output: html_document
---

# Setup
```{r}
pacman::p_load(
  tidyverse,    # Data manipulation and plotting
  here,         # Robust file paths
  cmdstanr,     # Stan interface
  rstan,
  posterior,    # Posterior analysis
  bayesplot,    # Plotting MCMC
  furrr,        # Parallel processing
  future,        # Future backend
  patchwork
)
```

```{r}
# Parallel Setup: Leave 2 cores free for system stability
n_cores <- parallel::detectCores() - 2
plan(multisession, workers = n_cores)
```

```{r}
# Optional: Set a default ggplot theme
theme_set(theme_classic())
```

# Simulating Data
```{r}
# Flag to control whether to regenerate simulation/fitting results for the knitted files
# Set to TRUE to rerun everything (takes time!), FALSE to load saved results.
regenerate_simulations <- TRUE # Or TRUE

```

# Define agents.
The two agents are: WBA and SBA

The variables:
ID : Participant ID
1_rating: The first rating the participant gives (0-8)
g_rating: The group rating feedback given after the first rating (0-8)
2_rating: The second rating the participant gives (0-8) <- this is what we want to predict
Feedback: How far is the g_rating from the 1_rating (-3 to 3)
Change: How much did the 2_rating change from the 1_rating (-7 to 7) but probably in the middle
Condition: Pre or Peri depending on whether the data was acquired before or during the pandemic

We can delete any rows with NAs!!

Social Conformity Experiment

 Goal: simulate agents that combine their own rating (FirstRating) with
       a group rating (GroupRating) to produce a second rating (SecondRating).

 Scale: ratings are integers 1–8 (1 = not trustworthy, 8 = very trustworthy)

 Key adaptation from the chapter's marble task:
   Marble task  → binary choice,   counts k out of n
   This task    → ordinal rating,  integers 1–8
   We convert ratings to pseudo-counts:
     k = rating - 1   (so rating 1 → 0 "blue", rating 8 → 7 "blue")
     n = 7            (total pseudo-marbles per source)
   This puts both sources on the same 0–7 count scale.
   Belief (theta) lives on [0, 1] and maps back:
     SecondRating = round(theta * 7) + 1
     
     
```{r}
library(tidyverse)
set.seed(42)
```

```{r}
 # Constants

# Rating scale boundaries
RATING_MIN <- 1
RATING_MAX <- 8
N_PSEUDO   <- 7   # pseudo sample size per source (= RATING_MAX - RATING_MIN)
 
# Jeffreys prior
PRIOR_ALPHA <- 0.5
PRIOR_BETA  <- 0.5
 
# Convert a 1–8 rating to a pseudo-count (0–7)
rating_to_count <- function(rating) as.integer(rating - 1)
 
# Convert a Beta belief (0–1) back to a 1–8 rating (rounded, clamped)
belief_to_rating <- function(theta) {
  raw <- round(theta * N_PSEUDO) + 1
  pmax(RATING_MIN, pmin(RATING_MAX, as.integer(raw)))
}
```

```{r}
# SECTION 2: Core belief-update function (shared by all agents)
#
# Implements Bayesian updating on a Beta distribution:
#   posterior = Beta(alpha_prior + w_self*k_self + w_group*k_group,
#                   beta_prior  + w_self*(n-k_self) + w_group*(n-k_group))
#
# Arguments:
#   rating_1  – participant's own rating (integer 1–8)
#   rating_g  – group's rating (integer 1–8)
#   w_self        – weight on own evidence  (WBA/PBA: free)
#   w_group       – weight on group evidence (WBA/PBA: free)

compute_posterior <- function(rating_1, rating_g,
                              w_self = 1, w_group = 1) {
  k_self  <- rating_to_count(rating_1)
  k_group <- rating_to_count(rating_g)
 
  alpha_post <- alpha_prior + w_self * k_self + w_group * k_group
  beta_post  <- beta_prior  + w_self * (N_PSEUDO - k_self) +
                               w_group * (N_PSEUDO - k_group)
 
  list(
    alpha = alpha_post,
    beta  = beta_post,
    expected_rate  = alpha_post / (alpha_post + beta_post)
  )
}
```

Three agent functions

 Each agent takes (first_rating, group_rating) and returns a SecondRating.
 All use probability matching:
   1. Sample theta ~ Beta(alpha_post, beta_post)   [belief about trustworthiness]
   2. SecondRating = round(theta * 7) + 1          [express as a 1–8 rating]

```{r}
# --- Agent B: Weighted Bayesian Agent (WBA) -----------------------------------
# Allows different weights for own vs. group evidence.
# w_self and w_group are both free parameters (>0).
#
# WARNING (from Chapter 10): the raw parameterisation is NOT identifiable —
# only the RATIO w_self/w_group matters. We keep raw weights here for
# simulation (they are interpretable), but the STAN model must reparameterise.
# See step 2 for the Stan solution.
 
wba_agent <- function(rating_1, rating_g, w_self, w_group) {
  
  # convert sample to count and compute posterior
  post <- compute_posterior(rating_1, rating_g, w_self, w_group)
  
  # sample from the posterior belief 
  theta <- rbeta(1, post$alpha, post$beta)
  
  #convert back to a rating
  belief_to_rating(theta) # Convert a Beta belief (0–1) back to a 1–8 rating (rounded, clamped)
}
```

```{r}
# --- Agent C: Proportional Bayesian Agent (PBA) ------------------------------
# Reparameterises WBA to a single identifiable weight w ∈ (0, 1):
#   w     = proportion of total weight given to own evidence
#   1 - w = proportion given to group evidence
#
# This fixes the identifiability problem:
#   w = 0.5  → equal weights (identical to SBA in belief direction,
#               but overall precision differs because total weight = 1, not 2)
#   w > 0.5  → trusts self more   (under-conformist)
#   w < 0.5  → trusts group more  (over-conformist / conformist)
 
pba_agent <- function(rating_1, rating_g, w) {
  # ensure weight is between 0 and 1
  stopifnot(w > 0, w < 1)
  
  # convert sample to count and compute posterior with proportional weights
  post  <- compute_posterior(rating_1, rating_g, w_self = w, w_group = 1 - w)
  
  # sample from the posterior belief
  theta <- rbeta(1, post$alpha, post$beta)
  
  # convert back to a rating
  belief_to_rating(theta)
}
```

# Simulate the task
```{r}
# Generates a tibble in the same format as the real datasets:
#   ID, FaceID, FirstRating, GroupRating, SecondRating
#
# Arguments:
#   agent_fn   – one of wba_agent / pba_agent (as a function)
#   agent_args – named list of extra args for the agent (e.g. list(w=0.7))
#   n_agents   – number of simulated participants
#   n_faces    – number of unique faces per participant
#   n_reps     – repetitions per (FirstRating, GroupRating) combination
#                (set n_faces=NULL to use full factorial design instead)

simulate_experiment <- function(agent_fn,
                                agent_args = list(),
                                n_agents   = 10,
                                n_faces    = 36,
                                n_reps     = 5) {
 

  # sample between 1-8 and generate a full factorial design of all combinations of first and group ratings
  
  
  # apply agents to the data
  
  
  # add ID and FaceID columns
  
  
  # return the simulated dataset

}


```

```{r}

```

3. Design 3–5 distinguishing scenarios
Pick parameter combinations that make SBA and WBA/PBA behave differently. E.g.:

Equal weights (SBA = WBA)
Strong self-trust (WBA with high w_self)
Strong conformity (WBA with high w_group)
Ignoring social info completely
Ignoring own prior completely
