---
title: "Script"
output: html_document
---

# Setup
```{r}
pacman::p_load(
  tidyverse,    # Data manipulation and plotting
  here,         # Robust file paths
  cmdstanr,     # Stan interface
  rstan,
  posterior,    # Posterior analysis
  bayesplot,    # Plotting MCMC
  furrr,        # Parallel processing
  future,        # Future backend
  patchwork,
  loo
)

main_path <- "/Users/magnusseverin/Documents/GitHub/Advanced_Cognitive_Modeling/Assignment03/src/"

setwd(main_path)

```

```{r}
# Parallel Setup: Leave 2 cores free for system stability
n_cores <- parallel::detectCores() - 2
plan(multisession, workers = n_cores)
```

```{r}
# Optional: Set a default ggplot theme
theme_set(theme_classic())
```

# Simulating Data
```{r}
# Flag to control whether to regenerate simulation/fitting results for the knitted files
# Set to TRUE to rerun everything (takes time!), FALSE to load saved results.
regenerate_simulations <- TRUE

```

# Define agents.
The two agents are: WBA and SBA

The variables:
ID : Participant ID
1_rating: The first rating the participant gives (0-8)
g_rating: The group rating feedback given after the first rating (0-8)
2_rating: The second rating the participant gives (0-8) <- this is what we want to predict
Feedback: How far is the g_rating from the 1_rating (-3 to 3)
Change: How much did the 2_rating change from the 1_rating (-7 to 7) but probably in the middle
Condition: Pre or Peri depending on whether the data was acquired before or during the pandemic

We can delete any rows with NAs!!

Social Conformity Experiment

 Goal: simulate agents that combine their own rating (FirstRating) with
       a group rating (GroupRating) to produce a second rating (SecondRating).

 Scale: ratings are integers 1–8 (1 = not trustworthy, 8 = very trustworthy)

 Key adaptation from the chapter's marble task:
   Marble task  → binary choice,   counts k out of n
   This task    → ordinal rating,  integers 1–8
   We convert ratings to pseudo-counts:
     k = rating - 1   (so rating 1 → 0 "blue", rating 8 → 7 "blue")
     n = 7            (total pseudo-marbles per source)
   This puts both sources on the same 0–7 count scale.
   Belief (theta) lives on [0, 1] and maps back:
     SecondRating = round(theta * 7) + 1
     
     
```{r}
library(tidyverse)
set.seed(42)
```

```{r}
# Rating scale boundaries
RATING_MIN <- 1
RATING_MAX <- 8
N_PSEUDO   <- 7   # pseudo sample size per source (= RATING_MAX - RATING_MIN)
 
# Jeffreys prior
alpha_prior <- 0.5
beta_prior <- 0.5
 

# Convert a Beta belief (0–1) back to a 1–8 rating (rounded, clamped)

```

```{r}
# SECTION 2: Core belief-update function (shared by all agents)
#
# Implements Bayesian updating on a Beta distribution:
#   posterior = Beta(alpha_prior + w_self*k_self + w_group*k_group,
#                   beta_prior  + w_self*(n-k_self) + w_group*(n-k_group))
#
# Arguments:
#   rating_1  – participant's own rating (integer 1–8)
#   rating_g  – group's rating (integer 1–8)
#   w_self        – weight on own evidence  (WBA/PBA: free)
#   w_group       – weight on group evidence (WBA/PBA: free)

compute_posterior <- function(rating_1, rating_g,
                              w_self = 1, w_group = 1) {
  k_self  <- rating_1 - 1 #converting ratings (1–8) to pseudo-counts (0–7)
  k_group <- rating_g - 1 
 
  alpha_post <- alpha_prior + w_self * k_self + w_group * k_group
  beta_post  <- beta_prior  + w_self * (N_PSEUDO - k_self) +
                               w_group * (N_PSEUDO - k_group)
 
  list(
    alpha = alpha_post,
    beta  = beta_post,
    expected_rate  = alpha_post / (alpha_post + beta_post)
  )
}
```

Three agent functions

 Each agent takes (first_rating, group_rating) and returns a SecondRating.
 All use probability matching:
   1. Sample theta ~ Beta(alpha_post, beta_post)   [belief about trustworthiness]
   2. SecondRating = round(theta * 7) + 1          [express as a 1–8 rating]

```{r}
# --- Agent B: Weighted Bayesian Agent (WBA) -----------------------------------
# Allows different weights for own vs. group evidence.
# w_self and w_group are both free parameters (>0).
 
wba_agent <- function(rating_1, rating_g, rho, kappa) {
  
  w_self <- rho * kappa 
  w_group <- (1 - rho) * kappa
  
  # convert sample to count and compute posterior
  post <- compute_posterior(rating_1, rating_g, w_self, w_group)
  
  # sample from the posterior belief 
  theta <- rbeta(1, post$alpha, post$beta)
  
  #convert back to a rating
  rating_2 <- round(theta * N_PSEUDO) + 1
  
  return(list(rating_1 = rating_1, rating_2 = rating_2, w_self = w_self, w_group = w_group))
}
```

```{r}
# --- Agent C: Proportional Bayesian Agent (PBA) ------------------------------
# Reparameterises WBA to a single identifiable weight w ∈ (0, 1):
#   w     = proportion of total weight given to own evidence
#   1 - w = proportion given to group evidence
#
# This fixes the identifiability problem:
#   w = 0.5  → equal weights (identical to SBA in belief direction,
#               but overall precision differs because total weight = 1, not 2)
#   w > 0.5  → trusts self more   (under-conformist)
#   w < 0.5  → trusts group more  (over-conformist / conformist)
 
pba_agent <- function(rating_1, rating_g, w) {
  
  # convert sample to count and compute posterior with proportional weights
  post  <- compute_posterior(rating_1, rating_g, w_self = w, w_group = 1 - w)
  
  # sample from the posterior belief
  theta <- rbeta(1, post$alpha, post$beta)
  
  # convert back to a rating
  rating_2 <- round(theta * N_PSEUDO) + 1
  
  return(list(rating_1 = rating_1, rating_2 = rating_2, w = w))
}
```

# Simulate the experiment 
```{r}
simulate_experiment <- function(agent_fn,
                                agent_args = list(),
                                n_faces    = 5) {
  
  offsets <- c(-3, -2, 2, 3)
  results <- vector("list", n_faces)
  idx <- 1
  
  for (face_id in 1:n_faces) {
      
      # Sample ratings
      rating_1 <- sample(1:8, 1)
      valid_offsets <- offsets[rating_1 + offsets >= 1 & rating_1 + offsets <= 8]
      feedback <- sample(valid_offsets, 1)
      rating_g <- rating_1 + feedback
      
      # Call agent
      result <- do.call(agent_fn, c(list(rating_1 = rating_1, rating_g = rating_g), agent_args))
      
      results[[idx]] <- tibble(
        FaceID       = face_id,
        FirstRating  = rating_1,
        GroupRating  = rating_g,
        SecondRating = result$rating_2,
        Feedback     = feedback,
        Change       = result$rating_2 - rating_1,
        total        = 8,
        N            = n_faces
      ) %>% bind_cols(as_tibble(agent_args))
      
      idx <- idx + 1
  } 
  
  bind_rows(results)
}
```

3. Design 3–5 distinguishing scenarios
Pick parameter combinations that make SBA and WBA/PBA behave differently. E.g.:

Equal weights (SBA = WBA)
Strong self-trust (WBA with high w_self)
Strong conformity (WBA with high w_group)
Ignoring social info completely
Ignoring own prior completely

Scenarios for WBA and PBA:
SCENARIO 01: (works only for wba - strong conformity)
kappa = 11
rho = 0.09
wd = 1
ws = 10

SCENARIO 02: (works for both models - balanced)
kappa = 1
rho = 0.5
wd = 0.5
ws= 0.5
 
SCENARIO 03: (works for both models - self focused)
kappa = 1
rho = 0.9
wd = 0.9
ws = 0.1



## WBA scenarios

### WBA: Strong conformity
```{r}
sim_df_wba_social <- simulate_experiment(wba_agent, agent_args = list(rho = 0.09, kappa = 11), n_faces = 100)
```

### WBA: Balanced
```{r}
sim_df_wba_balanced <- simulate_experiment(wba_agent, agent_args = list(rho = 0.5, kappa = 1), n_faces = 100)

```

### WBA: Self focused
```{r}
sim_df_wba_self <- simulate_experiment(wba_agent, agent_args = list(rho = 0.9, kappa = 1), n_faces = 100)

```

## PBA scenarios 

### PBA: Strong conformity...
```{r}
sim_df_pba_social <- simulate_experiment(pba_agent, agent_args = list(w = 0.1), n_faces = 100)

```

### PBA: Balanced
```{r}
sim_df_pba_balanced <- simulate_experiment(pba_agent, agent_args = list(w = 0.5), n_faces = 100)

```

### PBA: Self Focused
```{r}
sim_df_pba_self <- simulate_experiment(pba_agent, agent_args = list(w = 0.9), n_faces = 100)

```


```{r}
# combine and label your scenarios
sim_pba_all <- bind_rows(
  simulate_experiment(pba_agent, list(w = 0.1), n_faces = 200) %>% mutate(scenario = "Conformist (w=0.1)"),
  simulate_experiment(pba_agent, list(w = 0.5), n_faces = 200) %>% mutate(scenario = "Balanced (w=0.5)"),
  simulate_experiment(pba_agent, list(w = 0.9), n_faces = 200) %>% mutate(scenario = "Self-focused (w=0.9)")
)

# Then plot
ggplot(sim_pba_all, aes(x = Feedback, y = Change, color = scenario)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm") +
  facet_wrap(~scenario) +
  labs(title = "PBA: Change as a function of Group Feedback",
       x = "Feedback (GroupRating - FirstRating)",
       y = "Change (SecondRating - FirstRating)")
```

### Model Fitting 
```{r}
mod_pba <- cmdstan_model("ProbAgent.stan")


sampler_args <- list(
  chains = 4, 
  parallel_chains = 4, 
  iter_warmup = 1000, 
  iter_sampling = 2000, 
  refresh = 0
)
```

```{r}
mod_weighted_reparam <- cmdstan_model("WeightedAgent.stan")

sampler_args <- list(
  chains = 4, 
  parallel_chains = 4, 
  iter_warmup = 1000, 
  iter_sampling = 2000, 
  refresh = 0
)
```
##### Simulated data prep 
```{r}
stan_data_pba <- list(
  balanced = as.list(sim_df_pba_balanced),
  self     = as.list(sim_df_pba_self),
  social   = as.list(sim_df_pba_social)
)

stan_data_wba <- list(
  balanced = as.list(sim_df_wba_balanced),
  self     = as.list(sim_df_wba_self),
  social   = as.list(sim_df_wba_social)
)
```

```{r}
stan_data_pba$balanced$N <- as.integer(stan_data_pba$balanced$N[[1]])
stan_data_pba$self$N     <- as.integer(stan_data_pba$self$N[[1]])
stan_data_pba$social$N   <- as.integer(stan_data_pba$social$N[[1]])

stan_data_wba$balanced$N <- as.integer(stan_data_wba$balanced$N[[1]])
stan_data_wba$self$N     <- as.integer(stan_data_wba$self$N[[1]])
stan_data_wba$social$N   <- as.integer(stan_data_wba$social$N[[1]])
```

#### PBA Sim Model Fitting
```{r}
fit_pba_paths <- file.path(
  glue::glue("{main_path}/simmodels"),
  paste0("fit_pba_", names(stan_data_pba), ".rds")
) %>% set_names(names(stan_data_pba))

if (regenerate_simulations || !all(file.exists(fit_pba_paths))) {
  fit_pba <- imap(stan_data_pba, function(data, scenario) {
    fit <- mod_pba$sample(
      data          = data,
      chains        = 2,
      iter_warmup   = 500,
      iter_sampling = 2000,
      #fixed_param   = TRUE,
      refresh       = 0
    )
    fit$save_object(fit_pba_paths[[scenario]])
    fit
  })
} else {
  fit_pba <- map(fit_pba_paths, readRDS)
}
```

```{r}
readRDS("simmodels/fit_pba_self.rds")
```

#### WBA Sim Model Fitting 
```{r}
fit_wba_paths <- file.path(
  glue::glue("{main_path}/simmodels"),
  paste0("fit_wba_", names(stan_data_wba), ".rds")
) %>% set_names(names(stan_data_wba))

if (regenerate_simulations || !all(file.exists(fit_wba_paths))) {
  fit_wba <- imap(stan_data_wba, function(data, scenario) {
    fit <- mod_weighted_reparam$sample(
      data          = data,
      chains        = 2,
      iter_warmup   = 500,
      iter_sampling = 2000,
      #fixed_param   = TRUE,
      refresh       = 0
    )
    fit$save_object(fit_wba_paths[[scenario]])
    fit
  })
} else {
  fit_wba <- map(fit_wba_paths, readRDS)
}
```

```{r}
readRDS("simmodels/fit_wba_balanced.rds")
readRDS("simmodels/fit_wba_self.rds")
readRDS("simmodels/fit_wba_social.rds")

readRDS("simmodels/fit_pba_balanced.rds")
readRDS("simmodels/fit_pba_self.rds")
readRDS("simmodels/fit_pba_social.rds")
```

# Checking diagnostics
```{r}
# Check diagnostics
check_diagnostics <- function(fit, label) {
  diag_smry <- fit$diagnostic_summary()
  min_ebfmi <- min(diag_smry$ebfmi)
  
  smry <- fit$summary() %>% filter(!is.na(rhat))
  n_fail <- smry %>% filter(rhat > 1.01 | ess_bulk < 400 | ess_tail < 400) %>% nrow()
  
  cat(sprintf("%-25s | Divs: %3d | Min E-BFMI: %.2f | Failures: %d\n", 
              label, sum(diag_smry$num_divergent), min_ebfmi, n_fail))

}
cat("\n=== Weighted Bayesian Agent (Raw Parameterisation) ===\n")

iwalk(fit_wba, ~ check_diagnostics(.x, paste("WBA", .y)))
iwalk(fit_pba, ~ check_diagnostics(.x, paste("PBA", .y)))

```

```{r}
# WBA trace plots
imap(fit_wba, function(fit, scenario) {
  draws_df <- as_draws_df(fit$draws())
  
  p_trace <- mcmc_trace(draws_df, pars = c("rho", "kappa")) +
    theme_classic() +
    ggtitle(paste("WBA Trace Plot -", scenario))

  print(p_trace)
})


# PBA trace plots
imap(fit_pba, function(fit, scenario) {
  draws_df <- as_draws_df(fit$draws())
  
  p_trace <- mcmc_trace(draws_df, pars = c("w")) +
    theme_classic() +
    ggtitle(paste("PBA Trace Plot -", scenario))
  
  print(p_trace)
})
```

# Model Recovery 
```{r}
n_sims <- 20
plan(multisession, workers = parallel::detectCores() - 2)

if (regenerate_simulations || !file.exists("simmodels/ch10_model_recovery.rds")) {

    sim_grid <- tribble(~sim_id, ~true_model, ~w,   ~rho,  ~kappa, ~scenario,
  1, "PBA", 0.1, NA,   NA,  "PBA: Strong conformity",
  2, "PBA", 0.5, NA,   NA,  "PBA: Balanced",
  3, "PBA", 0.9, NA,   NA,  "PBA: Self focused",
  4, "WBA", NA,  0.09, 11,  "WBA: Strong conformity",
  5, "WBA", NA,  0.5,  1,   "WBA: Balanced",
  6, "WBA", NA,  0.9,  1,   "WBA: Self focused"
)

recovery_data <- sim_grid |>
  mutate(
    agent_data = pmap(list(true_model, w, rho, kappa), function(true_model, w, rho, kappa) {
      if (true_model == "PBA") {
        simulate_experiment(pba_agent, agent_args = list(w = w), n_faces = 10)
      } else {
        simulate_experiment(wba_agent, agent_args = list(rho = rho, kappa = kappa), n_faces = 10)
      }
    })
  )

  results_list <- list()
  
  recovery_data <- recovery_data[rep(seq_len(nrow(recovery_data)), each = 50), ]

  for (i in seq_len(nrow(recovery_data))) {
    id       <- recovery_data$sim_id[[i]]
    true_mod <- recovery_data$true_model[[i]]
    df       <- recovery_data$agent_data[[i]]
    
    # True parameters - derive wd/ws from whichever model generated the data
    if (true_mod == "PBA") {
      true_wd <- recovery_data$w[[i]]
      true_ws <- 1 - recovery_data$w[[i]]
    } else {
      true_wd <- recovery_data$rho[[i]] * recovery_data$kappa[[i]]
      true_ws <- (1 - recovery_data$rho[[i]]) * recovery_data$kappa[[i]]
    }
    
    # Stan data comes from the nested agent_data tibble
    stan_data <- list(
      N            = nrow(df),
      FirstRating  = df$FirstRating,
      SecondRating = df$SecondRating,
      GroupRating  = df$GroupRating,
      total        = df$total
    )
    

    # Fit models
    fit_p <- mod_pba$sample(data = stan_data,
                                     chains = 2, iter_warmup = 300,
                                     iter_sampling = 1300, refresh = 0,
                                     show_messages = FALSE, save_warmup = TRUE,
                                     output_dir = "simmodels/temp")

    fit_w <- mod_weighted_reparam$sample(data = stan_data,
                                 chains = 2, iter_warmup = 300,
                                 iter_sampling = 1300, refresh = 0,
                                 show_messages = FALSE, save_warmup = TRUE,
                                 output_dir = "simmodels/temp")

    loo_p <- fit_p$loo()
    loo_w <- fit_w$loo()

    weights <- loo::loo_model_weights(
      list(PBA = loo_p, WBA = loo_w),
      method = "stacking"
    )

    # Extract posterior means of effective weights (wd, ws) from each model
    pba_draws <- fit_p$draws(format = "df")
    wba_draws <- fit_w$draws(format = "df")

    results_list[[i]] <- tibble(
      sim_id     = id,
      scenario   = recovery_data$scenario[[i]], 
      true_model = true_mod,
      true_wd    = true_wd,
      true_ws    = true_ws,
      # Stacking weights
      weight_PBA = as.numeric(weights["PBA"]),
      weight_WBA = as.numeric(weights["WBA"]),
      # Posterior mean effective weights per candidate model
      pba_wd = mean(pba_draws$w),
      pba_ws = 1 - mean(pba_draws$w),
      wba_wd = mean(wba_draws$rho * wba_draws$kappa),
      wba_ws = mean((1 - wba_draws$rho) * wba_draws$kappa)
    )
  }

  recovery_results <- bind_rows(results_list)
  saveRDS(recovery_results, "simmodels/ch10_model_recovery.rds")

} else {
  recovery_results <- readRDS("simmodels/ch10_model_recovery.rds")
}
```
```{r}
# Model Recovery  #ATTEMPTED FIX 
n_sims <- 20
plan(multisession, workers = parallel::detectCores() - 2)
if (regenerate_simulations || !file.exists("simmodels/ch10_model_recovery.rds")) {
    sim_grid <- tribble(~sim_id, ~true_model, ~w,   ~rho,  ~kappa, ~scenario,
  1, "PBA", 0.1, NA,   NA,  "PBA: Strong conformity",
  2, "PBA", 0.5, NA,   NA,  "PBA: Balanced",
  3, "PBA", 0.9, NA,   NA,  "PBA: Self focused",
  4, "WBA", NA,  0.09, 11,  "WBA: Strong conformity",
  5, "WBA", NA,  0.5,  1,   "WBA: Balanced",
  6, "WBA", NA,  0.9,  1,   "WBA: Self focused"
)

  recovery_data <- sim_grid[rep(seq_len(nrow(sim_grid)), each = n_sims), ]  
  results_list <- list()
  
  for (i in seq_len(nrow(recovery_data))) {
    id       <- recovery_data$sim_id[[i]]
    true_mod <- recovery_data$true_model[[i]]
    
    # Simulate fresh data on every iteration
    df <- if (true_mod == "PBA") {
      simulate_experiment(pba_agent, agent_args = list(w = recovery_data$w[[i]]), n_faces = 10)
    } else {
      simulate_experiment(wba_agent, agent_args = list(rho = recovery_data$rho[[i]], kappa = recovery_data$kappa[[i]]), n_faces = 10)
    }
    
    # True parameters - derive wd/ws from whichever model generated the data
    if (true_mod == "PBA") {
      true_wd <- recovery_data$w[[i]]
      true_ws <- 1 - recovery_data$w[[i]]
    } else {
      true_wd <- recovery_data$rho[[i]] * recovery_data$kappa[[i]]
      true_ws <- (1 - recovery_data$rho[[i]]) * recovery_data$kappa[[i]]
    }
    
    # Stan data comes from freshly simulated df
    stan_data <- list(
      N            = nrow(df),
      FirstRating  = df$FirstRating,
      SecondRating = df$SecondRating,
      GroupRating  = df$GroupRating,
      total        = df$total
    )
    
    # Fit models
    fit_p <- mod_pba$sample(data = stan_data,
                                     chains = 2, iter_warmup = 300,
                                     iter_sampling = 1500, refresh = 0,
                                     show_messages = FALSE, save_warmup = TRUE,
                                     output_dir = "simmodels/temp")
    fit_w <- mod_weighted_reparam$sample(data = stan_data,
                                 chains = 2, iter_warmup = 300,
                                 iter_sampling = 1500, refresh = 0,
                                 show_messages = FALSE, save_warmup = TRUE,
                                 output_dir = "simmodels/temp")
    loo_p <- fit_p$loo()
    loo_w <- fit_w$loo()
    weights <- loo::loo_model_weights(
      list(PBA = loo_p, WBA = loo_w),
      method = "stacking"
    )
    pba_draws <- fit_p$draws(format = "df", inc_warmup = FALSE)
    wba_draws <- fit_w$draws(format = "df", inc_warmup = FALSE)
    results_list[[i]] <- tibble(
      sim_id     = id,
      scenario   = recovery_data$scenario[[i]], 
      true_model = true_mod,
      true_wd    = true_wd,
      true_ws    = true_ws,
      weight_PBA = as.numeric(weights["PBA"]),
      weight_WBA = as.numeric(weights["WBA"]),
      pba_wd = mean(pba_draws$w),
      pba_ws = mean(1 - pba_draws$w),
      wba_wd = mean(wba_draws$rho * wba_draws$kappa),
      wba_ws = mean((1 - wba_draws$rho) * wba_draws$kappa)
    )
  }
  recovery_results <- bind_rows(results_list)
  saveRDS(recovery_results, "simmodels/ch10_model_recovery.rds")
} else {
  recovery_results <- readRDS("simmodels/ch10_model_recovery.rds")
}
```
```{r}
loo::loo_compare(list(PBA = loo_p, WBA = loo_w))

```


```{r}
binary_matrix_scenario <- recovery_results |>
  pivot_longer(cols = starts_with("weight_"),
               names_to = "predicted_model", values_to = "weight") |>
  mutate(predicted_model = str_remove(predicted_model, "weight_")) |>
  group_by(sim_id, scenario) |>
  slice_max(weight, n = 1, with_ties = FALSE) |>
  ungroup() |>
  count(scenario, true_model, predicted_model) |>
  group_by(scenario) |>
  mutate(proportion = n / sum(n)) |>
  ungroup() |>
  mutate(
    true_model      = factor(true_model,      levels = c("PBA", "WBA")),
    predicted_model = factor(predicted_model, levels = c("PBA", "WBA"))
  ) |>
  complete(nesting(scenario, true_model), predicted_model, 
           fill = list(n = 0, proportion = 0))

mod_recov_bin <- binary_matrix_scenario |>
  ggplot(aes(x = predicted_model, y = true_model, fill = proportion)) +
  geom_tile(color = "white", linewidth = 1) +
  geom_text(aes(label = scales::percent(proportion, accuracy = 1)),
            size = 4, fontface = "bold") +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0, 1)) +
  scale_y_discrete(limits = rev) +
  facet_wrap(~ scenario) +
  labs(
    title = "Model Recovery: Binary Classification by Scenario",
    x = "Winning Candidate Model (Max Weight)",
    y = "True Generating Model",
    fill = "Win Rate"
  ) +
  theme(panel.grid = element_blank())

ggsave("/Users/magnusseverin/Documents/GitHub/Advanced_Cognitive_Modeling/Assignment03/src/mod_recov_bin.png", mod_recov_bin)

mod_recov_bin
```

```{r}
mod_recov_weight <- recovery_results |>
  group_by(scenario, true_model) |>
  summarise(
    Recovered_PBA = mean(weight_PBA),
    Recovered_WBA = mean(weight_WBA),
    .groups = "drop"
  ) |>
  pivot_longer(starts_with("Recovered"),
               names_to = "fitted_model", values_to = "weight") |>
  mutate(
    fitted_model = str_remove(fitted_model, "Recovered_"),
    fitted_model = factor(fitted_model, levels = c("PBA", "WBA")),
    true_model   = factor(true_model,   levels = c("PBA", "WBA"))
  ) |>
  ggplot(aes(x = fitted_model, y = true_model, fill = weight)) +
  geom_tile(color = "white", linewidth = 1) +
  geom_text(aes(label = scales::percent(weight, accuracy = 1)),
            size = 4, fontface = "bold", color = "white") +
  scale_fill_viridis_c(option = "mako", direction = -1, limits = c(0, 1)) +
  scale_y_discrete(limits = rev) +
  facet_wrap(~ scenario) +
  labs(
    title = "Model Recovery: Predictive Weight Allocation by Scenario",
    x = "Candidate Model", y = "True Generating Model",
    fill = "Avg. Weight"
  ) +
  theme(panel.grid = element_blank())

ggsave("mod_recov_weight.png", mod_recov_weight)

mod_recov_weight

```


```{r}
# Prior and posterior predictive checks
extract_predictive_data <- function(fit, stan_data, scenario_name, target_var) {
  # Extract draws for the target variable (prior_pred or posterior_pred)
  draws <- fit$draws(target_var, format = "matrix")
  
  # Flatten all draws into a long vector of predicted ratings
  predicted <- as.vector(draws)
  
  # Observed second ratings (repeated to match number of draw samples)
  n_draws <- nrow(draws)
  observed <- rep(stan_data$SecondRating, each = n_draws)
  
  tibble(
    value    = c(predicted, observed),
    type     = c(rep(paste(scenario_name, "Predicted"), length(predicted)),
                 rep("Observed", length(observed))),
    scenario = scenario_name
  )
}

plot_predictive_check <- function(data, title, subtitle) {
  data %>%
    ggplot(aes(x = value, fill = type, colour = type)) +
    geom_bar(position = "dodge", alpha = 0.6) +
    scale_x_continuous(breaks = 1:8) +
    labs(
      title    = title,
      subtitle = subtitle,
      x        = "Rating",
      y        = "Count",
      fill     = NULL,
      colour   = NULL
    ) +
    theme_minimal() +
    theme(plot.title = element_text(size = 9))
}

```

```{r}
# Helper function to generate and combine Prior vs. Posterior checks dynamically
generate_side_by_side_ppc <- function(fit_list, data_list, model_title) {
  
  # Map over all three scenarios (Self_Focused, Balanced, Socially_Influenced)
  plot_list <- imap(fit_list, function(fit, scenario) {
    
    # 1. Extract and plot Prior Predictive
    prior_data <- extract_predictive_data(
      fit = fit,
      stan_data = data_list[[scenario]],
      scenario_name = "Prior",
      target_var = "prior_pred"
    )
    
    p_prior <- plot_predictive_check(
      prior_data,
      title = paste(model_title, "-", str_replace_all(scenario, "_", " ")),
      subtitle = "Prior Predictive Check"
    ) + theme(legend.position = "none") # Hide individual legends to save space
    
    # 2. Extract and plot Posterior Predictive
    post_data <- extract_predictive_data(
      fit = fit,
      stan_data = data_list[[scenario]],
      scenario_name = "Posterior",
      target_var = "posterior_pred"
    )
    
    p_post <- plot_predictive_check(
      post_data,
      title = paste(model_title, "-", str_replace_all(scenario, "_", " ")),
      subtitle = "Posterior Predictive Check"
    ) + theme(axis.title.y = element_blank()) # Share Y axis visually with the prior plot
    
    # 3. Combine side-by-side
    return(p_prior | p_post)
  })
  
  # Stack the three scenarios vertically and collect the legend at the very bottom
  combined_all <- wrap_plots(plot_list, ncol = 1) + 
    plot_layout(guides = "collect") & 
    theme(legend.position = "bottom")
  
  return(combined_all)
}

cat("=== Proportional Bayesian Agent (PBA) Predictive Checks ===\n")
```


```{r}
# Helper function to generate and combine Prior vs. Posterior checks dynamically
generate_side_by_side_ppc <- function(fit_list, data_list, model_title) {
  
  # Map over all three scenarios (Self_Focused, Balanced, Socially_Influenced)
  plot_list <- imap(fit_list, function(fit, scenario) {
    
    # 1. Extract and plot Prior Predictive
    prior_data <- extract_predictive_data(
      fit = fit,
      stan_data = data_list[[scenario]],
      scenario_name = "Prior",
      target_var = "prior_pred"
    )
    
    p_prior <- plot_predictive_check(
      prior_data,
      title = paste(model_title, "-", str_replace_all(scenario, "_", " ")),
      subtitle = "Prior Predictive Check"
    ) + theme(legend.position = "none") # Hide individual legends to save space
    
    # 2. Extract and plot Posterior Predictive
    post_data <- extract_predictive_data(
      fit = fit,
      stan_data = data_list[[scenario]],
      scenario_name = "Posterior",
      target_var = "posterior_pred"
    )
    
    p_post <- plot_predictive_check(
      post_data,
      title = paste(model_title, "-", str_replace_all(scenario, "_", " ")),
      subtitle = "Posterior Predictive Check"
    ) + theme(axis.title.y = element_blank()) # Share Y axis visually with the prior plot
    
    # 3. Combine side-by-side
    return(p_prior | p_post)
  })
  
  # Stack the three scenarios vertically and collect the legend at the very bottom
  combined_all <- wrap_plots(plot_list, ncol = 1) + 
    plot_layout(guides = "collect") & 
    theme(legend.position = "bottom")
  
  return(combined_all)
}
```


```{r}
# Helper function to generate and combine Prior vs. Posterior checks dynamically
generate_side_by_side_ppc <- function(fit_list, data_list, model_title) {
  
  # Map over all three scenarios (Self_Focused, Balanced, Socially_Influenced)
  plot_list <- imap(fit_list, function(fit, scenario) {
    
    # 1. Extract and plot Prior Predictive
    prior_data <- extract_predictive_data(
      fit = fit,
      stan_data = data_list[[scenario]],
      scenario_name = "Prior",
      target_var = "prior_pred"
    )
    
    p_prior <- plot_predictive_check(
      prior_data,
      title = paste(model_title, "-", str_replace_all(scenario, "_", " ")),
      subtitle = "Prior Predictive Check"
    ) + theme(legend.position = "none") # Hide individual legends to save space
    
    # 2. Extract and plot Posterior Predictive
    post_data <- extract_predictive_data(
      fit = fit,
      stan_data = data_list[[scenario]],
      scenario_name = "Posterior",
      target_var = "posterior_pred"
    )
    
    p_post <- plot_predictive_check(
      post_data,
      title = paste(model_title, "-", str_replace_all(scenario, "_", " ")),
      subtitle = "Posterior Predictive Check"
    ) + theme(axis.title.y = element_blank()) # Share Y axis visually with the prior plot
    
    # 3. Combine side-by-side
    return(p_prior | p_post)
  })
  
  # Stack the three scenarios vertically and collect the legend at the very bottom
  combined_all <- wrap_plots(plot_list, ncol = 1) + 
    plot_layout(guides = "collect") & 
    theme(legend.position = "bottom")
  
  return(combined_all)
}

cat("=== Proportional Bayesian Agent (PBA) Predictive Checks ===\n")
```

```{r}
pba_combined <- generate_side_by_side_ppc(fit_pba, stan_data_pba, "PBA")
print(pba_combined)

wba_combined <- generate_side_by_side_ppc(fit_wba, stan_data_wba, "WBA")
print(wba_combined)

ggsave("pba_combined_ppc.png", pba_combined)
ggsave("wba_combined_ppc.png", wba_combined)
```

## Prior-posterior update checks
```{r}
# WBA Prior-Posterior Update (simulated)
wba_draws_balanced <- fit_wba$balanced$draws(c("rho", "kappa", "rho_prior", "kappa_prior"), format = "df")
wba_draws_self     <- fit_wba$self$draws(c("rho", "kappa", "rho_prior", "kappa_prior"), format = "df")
wba_draws_social   <- fit_wba$social$draws(c("rho", "kappa", "rho_prior", "kappa_prior"), format = "df")

# PLOT RHO
# wba self rho
p_rho_self <- ggplot(wba_draws_self) +
  geom_density(aes(x = rho_prior), fill = "red", colour = "black", alpha = 0.8) +
  geom_density(aes(x = rho), fill = "blue", colour = "black", alpha = 0.8) +
  geom_vline(xintercept = 0.9, linetype = "dashed", color = "black", linewidth = 1) +
  labs(title = "WBA rho - Self", x = "rho")

# wba balanced rho
p_rho_balanced <- ggplot(wba_draws_balanced) +
  geom_density(aes(x = rho_prior), fill = "red", colour = "black", alpha = 0.8) +
  geom_density(aes(x = rho), fill = "blue", colour = "black", alpha = 0.8) +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "black", linewidth = 1) +
  labs(title = "WBA rho - Balanced", x = "rho")

# wba social rho
p_rho_social <- ggplot(wba_draws_social) +
  geom_density(aes(x = rho_prior), fill = "red", colour = "black", alpha = 0.8) +
  geom_density(aes(x = rho), fill = "blue", colour = "black", alpha = 0.8) +
  geom_vline(xintercept = 0.09, linetype = "dashed", color = "black", linewidth = 1) +
  labs(title = "WBA rho - Social", x = "rho")


# PLOT KAPPA
# wba self kappa
p_kappa_self <- ggplot(wba_draws_self) +
  geom_density(aes(x = kappa_prior), fill = "red", colour = "black", alpha = 0.8) +
  geom_density(aes(x = kappa), fill = "blue", colour = "black", alpha = 0.8) +
  coord_cartesian(ylim = c(0, 0.5)) + 
  geom_vline(xintercept = 1, linetype = "dashed", color = "black", linewidth = 1) +
  labs(title = "WBA kappa - Self", x = "kappa")

# wba balanced kappa
p_kappa_balanced <- ggplot(wba_draws_balanced) +
  geom_density(aes(x = kappa_prior), fill = "red", colour = "black", alpha = 0.8) +
  geom_density(aes(x = kappa), fill = "blue", colour = "black", alpha = 0.8) +
  coord_cartesian(ylim = c(0, 0.5)) + 
  geom_vline(xintercept = 1, linetype = "dashed", color = "black", linewidth = 1) +
  labs(title = "WBA kappa - Balanced", x = "kappa")

# wba social kappa
p_kappa_social <- ggplot(wba_draws_social) +
  geom_density(aes(x = kappa_prior), fill = "red", colour = "black", alpha = 0.8) +
  geom_density(aes(x = kappa), fill = "blue", colour = "black", alpha = 0.8) +
  coord_cartesian(ylim = c(0, 0.5)) + 
  geom_vline(xintercept = 16, linetype = "dashed", color = "black", linewidth = 1) +
  labs(title = "WBA kappa - Social", x = "kappa")

# Combine
(p_rho_self | p_rho_balanced | p_rho_social) /
(p_kappa_self | p_kappa_balanced | p_kappa_social)
```


```{r}
# PBA Prior-Posterior Update (simulated)
pba_draws_balanced <- fit_pba$balanced$draws(c("w", "w_prior"), format = "df")
pba_draws_self     <- fit_pba$self$draws(c("w", "w_prior"), format = "df")
pba_draws_social   <- fit_pba$social$draws(c("w", "w_prior"), format = "df")

# PLOT W
# pba self w
pba_w_self <- ggplot(pba_draws_self) +
  geom_density(aes(x = w_prior), fill = "red", colour = "black", alpha = 0.8) +
  geom_density(aes(x = w), fill = "blue", colour = "black", alpha = 0.8) +
  geom_vline(xintercept = 0.9, linetype = "dashed", color = "black", linewidth = 1) +
  labs(title = "PBA w - Self", x = "w")

# pba balanced w
pba_w_balanced <- ggplot(pba_draws_balanced) +
  geom_density(aes(x = w_prior), fill = "red", colour = "black", alpha = 0.8) +
  geom_density(aes(x = w), fill = "blue", colour = "black", alpha = 0.8) +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "black", linewidth = 1) +
  labs(title = "PBA w - Balanced", x = "w")

# pba social w
pba_w_social <- ggplot(pba_draws_social) +
  geom_density(aes(x = w_prior), fill = "red", colour = "black", alpha = 0.8) +
  geom_density(aes(x = w), fill = "blue", colour = "black", alpha = 0.8) +
  geom_vline(xintercept = 0.1, linetype = "dashed", color = "black", linewidth = 1) +
  labs(title = "PBA w - Social", x = "w")


# Combine
(pba_w_self | pba_w_balanced | pba_w_social)


```

#### Experimental model fitting 
##### Exp Data Prep
```{r}
#We chose the cogsci df as our empirical data
cogsci_df <- read_csv("../data/cogsci_clean.csv")
```
#Fix Feedback = NA issue
```{r}
# Removing NA's
cogsci_df <- cogsci_df %>% filter(!is.na(Feedback))
```

```{r}
# Split into a named list, one dataframe per participant
cogsci_by_participant <- cogsci_df %>%
  group_by(ID) %>%
  group_split() %>%
  set_names(sort(unique(cogsci_df$ID)))

# Build stan data list for each participant
stan_data_exp <- map(cogsci_by_participant, function(df_p) {
  list(
    N        = nrow(df_p),
    FirstRating = as.integer(df_p$FirstRating),
    SecondRating = as.integer(df_p$SecondRating),
    GroupRating = as.integer(df_p$GroupRating),
    total  = rep(8L, nrow(df_p))
  )
})
```

### CogSci data exploration
```{r}

exploration_ratings_all_participants <- cogsci_df %>%
  pivot_longer(cols = c(FirstRating, SecondRating), 
               names_to = "Type", 
               values_to = "Value") %>%
  ggplot(aes(x = Value, fill = Type, color = Type)) +
  geom_histogram(position = "dodge") +
  scale_fill_manual(values = c("FirstRating" = "darkcyan", "SecondRating" = "coral")) +
  scale_color_manual(values = c("FirstRating" = "darkcyan", "SecondRating" = "coral")) +
  facet_wrap(~ID) +
  labs(
    title = "Ratings across all participants",
    x = "Rating",
    y = "Count"
  ) +
  theme_minimal()

exploration_ratings_all_participants
ggsave("exploration_ratings_all_participants.jpg", plot = exploration_ratings_all_participants, width = 12, height = 10, dpi = 300)



# change as function of group feedback (all participants)
exploration_change_all_participants <- ggplot(cogsci_df, aes(x = Feedback, y = Change)) +
  geom_point(alpha = 0.5, color = "darkcyan") +
  geom_smooth(method = "lm", se = TRUE, color = "darkcyan", fill = "darkcyan") +
  facet_wrap(~ID) +
  labs(
    title = "Change as a function of Group Feedback (all participants)",
    x = "Feedback (GroupRating - FirstRating)",
    y = "Change (SecondRating - FirstRating)"
  ) +
  theme_minimal()
```

```{r}

# change as function of group feedback (all participants)
exploration_change_all_participants <- ggplot(cogsci_df, aes(x = Feedback, y = Change)) +
  geom_point(alpha = 0.5, color = "darkcyan") +
  geom_smooth(method = "lm", se = TRUE, color = "darkcyan", fill = "darkcyan") +
  facet_wrap(~ID) +
  labs(
    title = "Change as a function of Group Feedback (all participants)",
    x = "Feedback (GroupRating - FirstRating)",
    y = "Change (SecondRating - FirstRating)"
  ) +
  theme_minimal()

exploration_change_all_participants
ggsave("exploration_change_all_participants.jpg", plot = exploration_change_all_participants, width = 12, height = 10, dpi = 300)
```

```{r}
# change as function of group feedback (one participant)
cogsci_df %>%
  filter(ID == 10) %>%
  ggplot(aes(x = Feedback, y = Change)) +
  geom_point(alpha = 0.5, color = "darkcyan") +
  geom_smooth(method = "lm", se = TRUE, color = "darkcyan", fill = "darkcyan") +
  labs(
    title = "Participant 1: Change as a function of Group Feedback",
    x = "Feedback (GroupRating - FirstRating)",
    y = "Change (SecondRating - FirstRating)"
  ) +
  theme_minimal()
```


#### PBA Experimental Model Fitting
```{r, include=FALSE}
fit_pba_paths <- file.path(
  glue::glue("{main_path}/simmodels"),
  paste0("fit_pba_", names(stan_data_exp), ".rds")
) %>% set_names(names(stan_data_exp))

if (regenerate_simulations || !all(file.exists(fit_pba_paths))) {
  fit_pba_exp <- imap(stan_data_exp, function(data, scenario) {
    fit_exp <- mod_pba$sample(
      data          = data,
      chains        = 2,
      iter_warmup   = 500,
      iter_sampling = 2000,
      #fixed_param   = TRUE,
      refresh       = 0
    )
    fit_exp$save_object(fit_pba_paths[[scenario]])
    fit_exp
  })
} else {
  fit_pba_exp <- map(fit_pba_paths, readRDS)
}
```


#### WBA Experimental Model Fitting 
```{r, include= FALSE}
fit_wba_paths <- file.path(
  glue::glue("{main_path}/simmodels"),
  paste0("fit_wba_", names(stan_data_exp), ".rds")
) %>% set_names(names(stan_data_exp))

if (regenerate_simulations || !all(file.exists(fit_wba_paths))) {
  fit_wba_exp <- imap(stan_data_exp, function(data, scenario) {
    fit_exp <- mod_weighted_reparam$sample(
      data          = data,
      chains        = 2,
      iter_warmup   = 500,
      iter_sampling = 2000,
      #fixed_param   = TRUE,
      refresh       = 0
    )
    fit_exp$save_object(fit_wba_paths[[scenario]])
    fit_exp
  })
} else {
  fit_wba_exp <- map(fit_wba_paths, readRDS)
}
```


#### Extract per-participant estimates
```{r}
# WBA: one rho per participant
wba_estimates <- imap_dfr(fit_wba_exp, function(fit_exp, pid) {
  fit_exp$summary(c("rho", "kappa")) %>% mutate(ID = pid)
})

# PBA: one w per participant  
pba_estimates <- imap_dfr(fit_pba_exp, function(fit_exp, pid) {
  fit_exp$summary("w") %>% mutate(ID = pid)
})

#Just to verify that it includes the participants
wba_estimates <- wba_estimates %>% mutate(ID = as.integer(ID))
pba_estimates <- pba_estimates %>% mutate(ID = as.integer(ID))

wba_estimates
pba_estimates
```

```{r}
# example participant
readRDS("simmodels/fit_pba_1.rds")
```

# Data exploration
```{r}
# WBA: rho
p_rho <- wba_estimates %>%
  filter(variable == "rho") %>%
  arrange(mean) %>%
  mutate(ID = factor(ID, levels = ID)) %>%
  ggplot(aes(x = ID, y = mean)) +
  geom_point(size = 2, color = "#4CAF50") +
  labs(title = "WBA: rho per participant",
       x = "Participant", y = "rho") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# WBA: kappa
p_kappa <- wba_estimates %>%
  filter(variable == "kappa") %>%
  arrange(mean) %>%
  mutate(ID = factor(ID, levels = ID)) %>%
  ggplot(aes(x = ID, y = mean)) +
  geom_point(size = 2, color = "#E91E63") +
  labs(title = "WBA: Kappa per participant",
       x = "Participant", y = "kappa") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# PBA: w
p_w <- pba_estimates %>%
  filter(variable == "w") %>%
  arrange(mean) %>%
  mutate(ID = factor(ID, levels = ID)) %>%
  ggplot(aes(x = ID, y = mean)) +
  geom_point(size = 2,color = "#2196F3") +
  labs(title = "PBA: weight per participant",
       x = "Participant", y = "w") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

p_rho / p_kappa / p_w
```

```{r}
# PBA: w per participant
p_w <- pba_estimates %>%
  filter(variable == "w") %>%
  ggplot(aes(x = reorder(ID, mean), y = mean)) +
  geom_col(fill = "steelblue") +
  #geom_hline(aes(yintercept = mean(mean)), linewidth = 0.8) +
  #geom_hline(yintercept = 0.5, linetype = "dashed") +
  labs(title = "PBA: w", x = "Participants", y = "w") +
  theme_bw() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())

# WBA: rho per participant
p_rho <- wba_estimates %>%
  filter(variable == "rho") %>%
  ggplot(aes(x = reorder(ID, mean), y = mean)) +
  geom_col(fill = "orange") +
  #geom_hline(aes(yintercept = mean(mean)), linewidth = 0.8) +
  #geom_hline(yintercept = 0.5, linetype = "dashed") +
  labs(title = "WBA: rho", x = "Participants", y = "rho") +
  theme_bw() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())

# WBA: kappa per participant
p_kappa <- wba_estimates %>%
  filter(variable == "kappa") %>%
  ggplot(aes(x = reorder(ID, mean), y = mean)) +
  geom_col(fill = "coral") +
  #geom_hline(aes(yintercept = mean(mean)), linewidth = 0.8) +
  labs(title = "WBA: kappa", x = "Participants", y = "kappa") +
  theme_bw() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())

(p_rho | p_kappa )/ p_w
```

# Checking diagnostics experimental

```{r}
# Check diagnostics
check_diagnostics <- function(fit_exp, label) {
  diag_smry <- fit_exp$diagnostic_summary()
  min_ebfmi <- min(diag_smry$ebfmi)
  
  smry <- fit_exp$summary() %>% filter(!is.na(rhat))
  n_fail <- smry %>% filter(rhat > 1.01 | ess_bulk < 400 | ess_tail < 400) %>% nrow()
  
  cat(sprintf("%-25s | Divs: %3d | Min E-BFMI: %.2f | Failures: %d\n", 
              label, sum(diag_smry$num_divergent), min_ebfmi, n_fail))

}
cat("\n=== Weighted Bayesian Agent (Raw Parameterisation) ===\n")

iwalk(fit_wba_exp, ~ check_diagnostics(.x, paste("WBA", .y)))
iwalk(fit_pba_exp, ~ check_diagnostics(.x, paste("PBA", .y)))

```



```{r}
# WBA trace plots - cogsci dataset
imap(fit_wba_exp, function(fit_exp, scenario) {
  draws_df <- as_draws_df(fit_exp$draws())
  
  p_trace <- mcmc_trace(draws_df, pars = c("rho", "kappa")) +
    theme_classic() +
    ggtitle(paste("WBA Trace Plot -", scenario))

    ggsave(
    filename = paste0("trace_wba_", scenario, ".png"),
    plot = p_trace,
    width = 10, height = 6, dpi = 300
  )
})
```

```{r}
# PBA trace plots - cogsci dataset
imap(fit_pba_exp, function(fit_exp, scenario) {
  draws_df <- as_draws_df(fit_exp$draws())
  
  p_trace <- mcmc_trace(draws_df, pars = c("w")) +
    theme_classic() +
    ggtitle(paste("PBA Trace Plot -", scenario))
  
  #print(p_trace)
})
```

## Prior-posterior update checks experimental
```{r}
# WBA Prior-Posterior Update (experimental)
invisible(imap(fit_wba_exp, function(fit, pid) {
  draws <- fit$draws(c("rho", "kappa", "rho_prior", "kappa_prior"), format = "df")
  
  p_rho <- ggplot(draws) +
    geom_density(aes(x = rho_prior), fill = "red", colour = "black", alpha = 0.8) +
    geom_density(aes(x = rho), fill = "blue", colour = "black", alpha = 0.8) +
    labs(title = paste("WBA rho - Participant", pid), x = "rho")
  
  p_kappa <- ggplot(draws) +
    geom_density(aes(x = kappa_prior), fill = "red", colour = "black", alpha = 0.8) +
    geom_density(aes(x = kappa), fill = "blue", colour = "black", alpha = 0.8) +
    labs(title = paste("WBA kappa - Participant", pid), x = "kappa")
  
  print(p_rho / p_kappa)
}))

```


```{r}
# PBA Prior-Posterior Update (experimental)
invisible(imap(fit_pba_exp, function(fit, pid) {
  draws <- fit$draws(c("w", "w_prior"), format = "df")
  
  p_w <- ggplot(draws) +
    geom_density(aes(x = w_prior), fill = "red", colour = "black", alpha = 0.8) +
    geom_density(aes(x = w), fill = "blue", colour = "black", alpha = 0.8) +
    labs(title = paste("PBA w - Participant", pid), x = "w")
  
  print(p_w)
}))
```

## Prior and Posterior predictive checks

```{r}
# Prior and posterior predictive checks
extract_predictive_data_exp <- function(fit_exp, stan_data_exp, scenario_name, target_var) {
  # Extract draws for the target variable (prior_pred or posterior_pred)
  draws_exp <- fit_exp$draws(target_var, format = "matrix")
  
  # Flatten all draws into a long vector of predicted ratings
  predicted <- as.vector(draws_exp)
  
  # Observed second ratings (repeated to match number of draw samples)
  n_draws <- nrow(draws_exp)
  observed <- rep(stan_data_exp$SecondRating, times = n_draws)
  
  tibble(
    value    = c(predicted, observed),
    type     = c(rep(paste(scenario_name, "Predicted"), length(predicted)),
                 rep("Observed", length(observed))),
    scenario = scenario_name
  )
}

plot_predictive_check <- function(data, title, subtitle) {
  data %>%
    ggplot(aes(x = value, fill = type, colour = type)) +
    geom_bar(position = "dodge", alpha = 0.6) +
    scale_x_continuous(breaks = 1:8) +
    labs(
      title    = title,
      subtitle = subtitle,
      x        = "Rating",
      y        = "Count",
      fill     = NULL,
      colour   = NULL
    ) +
    theme_minimal() +
    theme(plot.title = element_text(size = 9))
}
```


```{r}
# Helper function to generate and combine Prior vs. Posterior checks dynamically
generate_side_by_side_ppc_exp <- function(fit_list, data_list, model_title) {
  
  # Map over all three scenarios (Self_Focused, Balanced, Socially_Influenced)
  plot_list <- imap(fit_list, function(fit, scenario) {
    
    # 1. Extract and plot Prior Predictive
    prior_data <- extract_predictive_data_exp(
      fit_exp = fit,
      stan_data_exp = data_list[[scenario]],
      scenario_name = "Prior",
      target_var = "prior_pred"
    )
    
    p_prior <- plot_predictive_check(
      prior_data,
      title = paste(model_title, "-", str_replace_all(scenario, "_", " ")),
      subtitle = "Prior Predictive Check"
    ) + theme(legend.position = "none") # Hide individual legends to save space
    
    # 2. Extract and plot Posterior Predictive
    post_data <- extract_predictive_data_exp(
      fit_exp = fit,
      stan_data_exp = data_list[[scenario]],
      scenario_name = "Posterior",
      target_var = "posterior_pred"
    )
    
    p_post <- plot_predictive_check(
      post_data,
      title = paste(model_title, "-", str_replace_all(scenario, "_", " ")),
      subtitle = "Posterior Predictive Check"
    ) + theme(axis.title.y = element_blank()) # Share Y axis visually with the prior plot
    
    # 3. Combine side-by-side
    return(p_prior | p_post)
  })
  
  # Stack the three scenarios vertically and collect the legend at the very bottom
  combined_all <- wrap_plots(plot_list, ncol = 1) + 
    plot_layout(guides = "collect") & 
    theme(legend.position = "bottom")
  
  return(combined_all)
}

cat("=== Proportional Bayesian Agent (PBA) Predictive Checks ===\n")
```


```{r}
pid <- "10"

pba_single <- generate_side_by_side_ppc_exp(
  fit_list   = fit_pba_exp[pid],
  data_list  = stan_data_exp[pid],
  model_title = paste("PBA - Participant", pid)
)
print(pba_single)

wba_single <- generate_side_by_side_ppc_exp(
  fit_list   = fit_wba_exp[pid],
  data_list  = stan_data_exp[pid],
  model_title = paste("WBA - Participant", pid)
)
print(wba_single)

ppc_combined <- pba_single / wba_single

ggsave(
  filename = "id10_ppc_plot.png",
  plot = ppc_combined
)
```

```{r}
# Get unique participant IDs from the data
participant_ids <- unique(cogsci_df$ID)

agent_loo <- map(participant_ids, function(id) {
  
  # Load both model fits for this participant
  fit_pba <- readRDS(paste0("simmodels/fit_pba_", id, ".rds"))
  fit_wba <- readRDS(paste0("simmodels/fit_wba_", id, ".rds"))
  
  # Compute LOO for each model
  # This uses the log_lik vector declared in generated quantities
  loo_pba <- fit_pba$loo()
  loo_wba <- fit_wba$loo()
  
  # Sum pointwise ELPD across all trials to get one number per agent per model
  # Higher = better out-of-sample predictive accuracy
  tibble(
    agent_id   = id,
    elpd_PBA   = sum(loo_pba$pointwise[, "elpd_loo"]),
    elpd_WBA   = sum(loo_wba$pointwise[, "elpd_loo"]),
    # Winning model = whichever has higher ELPD for this agent
    best_model = if_else(elpd_PBA > elpd_WBA, "PBA", "WBA")
  )
  
}) |> list_rbind()
```

#### Another model comparison
```{r}
# Pool all trial-level log_lik across participants into one matrix
# by stacking rows (trials) from each participant
log_lik_pba <- do.call(rbind, map(fit_pba_exp, ~ .x$draws("log_lik", format = "matrix")))
log_lik_wba <- do.call(rbind, map(fit_wba_exp, ~ .x$draws("log_lik", format = "matrix")))

# Compute population-level LOO
loo_pba <- loo(log_lik_pba)
loo_wba <- loo(log_lik_wba)

# Population-level comparison
cat("=== Population-Level Comparison (Experimental Data) ===\n")
print(loo::loo_compare(list("PBA" = loo_pba, "WBA" = loo_wba)))

# Stacking weights
cat("\nStacking weights:\n")
print(round(loo_model_weights(list("PBA" = loo_pba, "WBA" = loo_wba), method = "stacking"), 3))

```

```{r}
# Pool all trial-level log_lik across participants into one matrix
log_lik_pba <- do.call(rbind, map(fit_pba_exp, ~ .x$draws("log_lik", format = "matrix")))
log_lik_wba <- do.call(rbind, map(fit_wba_exp, ~ .x$draws("log_lik", format = "matrix")))

# Compute population-level LOO
loo_pba <- loo(log_lik_pba)
loo_wba <- loo(log_lik_wba)

# Population-level comparison
cat("=== Population-Level Comparison (Experimental Data) ===\n")
print(loo::loo_compare(list("PBA" = loo_pba, "WBA" = loo_wba)))

# Stacking weights
cat("\nStacking weights:\n")
print(round(loo::loo_model_weights(list("PBA" = loo_pba, "WBA" = loo_wba), method = "stacking"), 3))
```

```{r}
agent_elpd <- imap_dfr(fit_pba_exp, function(fit_pba, pid) {
  fit_wba <- fit_wba_exp[[pid]]
  
  # Get pointwise ELPD for each model for this participant
  loo_pba <- fit_pba$loo()
  loo_wba <- fit_wba$loo()
  
  tibble(
    agent_id = as.integer(pid),
    elpd_PBA = sum(loo_pba$pointwise[, "elpd_loo"]),
    elpd_WBA = sum(loo_wba$pointwise[, "elpd_loo"])
  )
}) |>
  mutate(
    best_model     = if_else(elpd_PBA > elpd_WBA, "PBA", "WBA"),
    # ELPD advantage of WBA over PBA (positive = WBA better)
    elpd_advantage = elpd_WBA - elpd_PBA
  )

```

```{r}
# Left: per-agent ELPD advantage
p_agent_advantage <- ggplot(agent_elpd,
       aes(x = reorder(agent_id, elpd_advantage),
           y = elpd_advantage,
           fill = elpd_advantage > 0)) +
  geom_col(alpha = 0.8) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
  scale_fill_manual(values = c("TRUE" = "#984ea3", "FALSE" = "#4daf4a"),
                    labels = c("TRUE" = "WBA better", "FALSE" = "PBA better")) +
  coord_flip() +
  labs(
    title    = "Per-Agent ELPD: WBA vs. PBA",
    subtitle = "Positive = WBA predicts better. Negative = PBA predicts better.",
    x        = "Participant (sorted by advantage)",
    y        = "ELPD Advantage (WBA − PBA)",
    fill     = "Better Model"
  ) +
  theme(
    axis.text.y  = element_text(size = 6),
    legend.position = "bottom"
  )

# Right: scatter of per-agent PBA vs WBA ELPD
p_agent_scatter <- ggplot(agent_elpd,
       aes(x = elpd_PBA, y = elpd_WBA,
           shape = best_model, colour = best_model)) +
  geom_abline(linetype = "dashed", colour = "grey50") +
  geom_point(size = 3, alpha = 0.8) +
  scale_colour_manual(values = c("PBA" = "#4daf4a", "WBA" = "#984ea3")) +
  scale_shape_manual(values = c("PBA" = 16, "WBA" = 17)) +
  labs(
    title    = "Agent-Level LOO: PBA vs. WBA",
    subtitle = "Points above diagonal: WBA predicts better. Shape = winning model.",
    x        = "Agent ELPD under PBA",
    y        = "Agent ELPD under WBA",
    colour   = "Best Model", shape = "Best Model"
  ) +
  theme(legend.position = "bottom")

ggsave("mod_comp_adv.png", p_agent_advantage)
ggsave("mod_comp_scatter.png", p_agent_scatter)

p_agent_advantage | p_agent_scatter
```

